### 2025-03-28 analysis/hypothesis.qmd

---
title: "Hypotheses"
format: 
  html:
    code-tools: false
---

## Primary Hypothesis

We hypothesize that:

- **H1:** Higher levels of procrastination are associated with:
  - Increased risk of transitioning from Normal Cognition to MCI $(\beta = x, p < .05)$
  - Increased risk of transitioning from MCI to Dementia $(\beta = y, p < .05)$

- **H0:** There is no association between procrastination and cognitive decline transitions

In order to test this hypothesis we will be using a **discrete time Markov Model**