### 2025-03-28 index.qmd

---
title: "Procrastination and Dementia"
author: 
  - name: "Cormac Monaghan"
    url: "https://c-monaghan.github.io/"
    orcid: "https://orcid.org/0000-0001-9012-3060"
    affiliation:
      - "Hamilton Institute, Maynooth University, Ireland"
      - "Department of Psychology, Maynooth University, Ireland"
  - name: "Rafael de Andrade Moral"
    url: "https://rafamoral.github.io/"
    orcid: "https://orcid.org/0000-0002-0875-3563"
    affiliation:
      - "Department of Mathematics and Statistics, Maynooth University, Ireland"
  - name: "Joanna McHugh Power"
    orcid: "https://orcid.org/0000-0002-7387-3107"
    affiliation: 
      - "Department of Psychology, Maynooth University, Ireland"
format: 
  html:
    toc: false
sidebar: false
---

### Abstract

**This will come eventually**