### 2025-03-28 analysis/continous-markov.qmd
Updating site

---
title: "Continous Time Markov Model"
format: html
---

# Setting up

## Loading packages

```{r packages}
#| code-fold: true
#| code-summary: "Check our my packages"
#| warning: false
#| message: false

library(dplyr)
library(tidyr)
library(ggplot2)
```

## Loading functions

```{r functions}
#| code-fold: true
#| code-summary: "Check our my functions"

extract_years <- function(data, years, impute = TRUE, cog_total = FALSE, absorbing = TRUE) {
  # Extracts cognitive function data for specified years from a dataset.
  # Converts numeric cognitive status codes (1, 2, 3) into descriptive labels
  # ("Normal Cognition", "MCI", "Dementia") for easier interpretation.
  # Arguments:
  #   - data: The input dataset containing cognitive function data.
  #   - years: A vector of years for which data should be extracted.
  # Returns:
  #   - A dataset with ID and cognitive status columns for the specified years.

  # Create dynamic column names based on the years provided
  cogfunction_cols <- paste0("cogfunction", years)
  cogtotal_cols    <- paste0("cogtot27_imp", years)

  if(cog_total == FALSE) {
  data <- data |>
    # Select only the ID column and cognitive function columns for the specified years
    select(ID, any_of(cogfunction_cols)) |>
    mutate(across(!c(ID), ~ case_when(
      .x == 1 ~ "Normal Cognition",
      .x == 2 ~ "MCI",
      .x == 3 ~ "Dementia",
      TRUE ~ NA_character_  # To handle missing/other cases
    )))
  } else {
    data <- data |>
      # Select only the ID column and cognitive function columns for the specified years
      select(ID, any_of(cogtotal_cols)) |>
      rename_with( .cols = !ID, .fn = ~ stringr::str_replace(
        string = .x,
        pattern = "cogtot27_imp", 
        replacement = "cog_score_"))
  }

  if(impute == TRUE){
    data <- data |>
      tidyr::pivot_longer(
        cols = !ID,
        names_to = "Wave",
        values_to = "Status") |>
      group_by(ID) |>
      tidyr::fill(Status, .direction = "down") |>
      ungroup() |>
      tidyr::pivot_wider(names_from = "Wave", values_from = "Status")
  }

  if(absorbing == TRUE) {
    data <- data |>
      tidyr::pivot_longer(
        cols = !ID,
        names_to = "Wave",
        values_to = "Status") |>
      group_by(ID) |>
      mutate(Status = ifelse(cumany(Status == "Dementia"), "Dementia", Status)) |>
      ungroup() |>
      tidyr::pivot_wider(names_from = "Wave", values_from = "Status")
  }

  return(data)
}

create_transitions <- function(data, absorbing = FALSE){
  # Reshapes data from wide to long format to track cognitive status transitions over time.
  # Calculates the next wave's cognitive status for each individual and creates a transition column.
  # Optionally treats "Dementia" as an absorbing state, meaning once an individual is classified
  # with dementia, their status cannot change in subsequent waves.
  # Arguments:
  #   - data: The dataset containing cognitive status data.
  #   - absorbing: A logical flag indicating whether "Dementia" should be treated 
  #   as an absorbing state.
  # Returns:
  #   - A dataset with transition information, including current and next wave statuses.

  # Reshape the data from wide to long format to track cognitive status over waves
  data <- data |>
    select(ID, starts_with("cogfunction")) |>
    tidyr::pivot_longer(cols = !ID,
                        names_to = "Wave",
                        values_to = "Status") |>
    mutate(Wave = as.factor(stringr::str_replace(Wave, "cogfunction", ""))) |>
    # Arrange by ID and Wave to prepare for transition calculation
    arrange(ID, Wave) |>
    group_by(ID) |>
    # Get the next wave's cognitive status for each person
    mutate(next_wave_status = lead(Status)) |>
    ungroup()

  # We can optionally specify dementia as an absorbing state
  # Once an individual is classified with dementia they cannot be classified
  # with anything else 
  if(absorbing == TRUE) {
    data <- data |>
      group_by(ID) |>
      mutate(
        Status = ifelse(cumany(Status == "Dementia"), "Dementia", Status),
        next_wave_status = ifelse(cumany(Status == "Dementia"), "Dementia", next_wave_status),
        transition = paste(Status, next_wave_status, sep = " to ")
      ) |>
      filter(Wave %in% c(2016, 2018, 2020))
  }

  # Filter out rows where either the current or next status is missing
  data <- data |>
    group_by(ID) |>
    filter(!is.na(Status), !is.na(next_wave_status)) |>
    ungroup() |>
    # Create a new column representing the transition from one status to the next
    mutate(
      transition = paste(Status, next_wave_status, sep = " to "),
      transition = factor(
        transition, 
        levels = c("Normal Cognition to Normal Cognition", "Normal Cognition to MCI",
                   "Normal Cognition to Dementia", "MCI to Normal Cognition", "MCI to MCI",
                   "MCI to Dementia", "Dementia to Dementia") 
        ))

  return(data)
}

calculate_probabilties <- function(data) {
  # Calculates the proportion of each cognitive status transition in the dataset.
  # Counts the occurrences of each transition and computes their relative probabilities.
  # Arguments:
  #   - data: The dataset containing transition information.
  # Returns:
  #   - A dataset with transition probabilities, split into "from" and "to" states.

  # Calculate the proportion of each transition by dividing by the total count
  data |>
    count(transition) |>
    mutate(prop = n / sum(n)) |>
    tidyr::separate(transition, into = c("from", "to"), sep = " to ")
}

transition_matrix <- function(data, longitudinal = FALSE) {
  # Constructs a transition matrix from the calculated transition probabilities.
  # The matrix represents the probability of moving from one cognitive state to another.
  # Arguments:
  #   - data: The dataset containing transition probabilities.
  #   - longitudinal: A TRUE/FALSE statement indicating if the probabilities are from multiple waves
  # Returns:
  #   - A transition matrix with rows representing "from" states and columns 
  #   representing "to" states.

  if(longitudinal == FALSE) {
    # Defining empty matrix matrix
    states <- c("Normal Cognition", "MCI", "Dementia")

    transition_matrix <- matrix(
      0,
      nrow = length(states),
      ncol = length(states),
      dimnames = list(from_state = states, to_state = states))

    # Fill the transition matrix with probabilities
    for (i in 1:nrow(data)) {
      from <- data$from[i]
      to <- data$to[i]
      prob <- data$prop[i]

      transition_matrix[from, to] <- prob
    }
  } else if(longitudinal == TRUE){

    # Aggregate by 'from' and 'to' states to sum transition probabilities
    data_agg <- data |>
      group_by(from, to) |>
      summarise(transition_prob = sum(prop), .groups = "drop")

    transition_matrix <- data_agg |>
      tidyr::pivot_wider(names_from = to, values_from = transition_prob, values_fill = 0) |>
      tibble::column_to_rownames("from") |>
      as.matrix()
  }

  return(transition_matrix)
}

plot_matrix <- function(data, year, ts = FALSE) {
  # Visualizes the transition matrix as a heatmap.
  # Uses color gradients to represent transition probabilities and includes labels for clarity.
  # Arguments:
  #   - data: The transition matrix or dataset to plot.
  #   - year: The year or time period associated with the data.
  #   - ts: A logical flag indicating whether the title should include a time series label.
  # Returns:
  #   - A ggplot object representing the transition probabilities as a heatmap.

  if(ts == TRUE) {
    title <- paste0("Transition Probabilities Across Cognitive States - ", year)
  } else {
    title <- paste0("Transition Probabilities Across Cognitive States (", year, "-2022)")
  }

  # Reshaping data for plotting
  data <- data |>
    as.data.frame(row.names = FALSE) |>
    mutate(from_state = c("Normal Cognition", "MCI", "Dementia")) |>
    reshape2::melt(id.vars = "from_state", variable.name = "to_state", value.name = "probability")

  # Plotting probabilities
  data |>
    ggplot(aes(x = from_state, y = to_state, fill = probability)) +
    geom_tile() +
    scale_fill_viridis_c(alpha = 0.8) +
    labs(title = title,
         x = "To", y = "From", fill = "Prob") +
    geom_text(aes(label = round(probability, digits = 3)), size = 5) +
    theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),
          axis.title = element_text(size = 10),
          axis.text = element_text(size = 8),
          panel.grid = element_blank()
    )
}

normalise <- function(x) {
  # Normalizes a matrix so that each row sums to 1.
  # Arguments:
  #   - x: The input matrix to normalize.
  # Returns:
  #   - A normalized matrix where each row sums to 1.
  x / rowSums(x)
}

reshape_matrix <- function(matrix, wave) {
  # Reshapes a transition matrix into a long format for easier analysis and visualization.
  # Adds a "Wave" column to identify the time period associated with the matrix.
  # Arguments:
  #   - matrix: The transition matrix to reshape.
  #   - wave: The wave or time period associated with the matrix.
  # Returns:
  #   - A long-format dataset with "from", "to", and "transition_prob" columns.

  matrix |>
    as.data.frame() |>
    tibble::rownames_to_column(var = "from") |>
    tidyr::pivot_longer(cols = -from, names_to = "to", values_to = "transition_prob") |>
    mutate(Wave = wave)
}

plot_transition_wave <- function(data, wave) {
  # Plots transition probabilities for a specific wave as a heatmap.
  # Filters the data to include only the main transitions and adds labels for clarity.
  # Arguments:
  #   - data: The dataset containing transition probabilities.
  #   - wave: The wave or time period to plot.
  # Returns:
  #   - A ggplot object representing the transition probabilities for the 
  #   specified wave.

  # Main transitions per wave
  filter_data <- data.frame(
    from = c("Normal Cognition", "Normal Cognition", "Normal Cognition",  "MCI", "MCI", "MCI", "Dementia", "Dementia", "Dementia"),
    to = c("Normal Cognition", "MCI", "Dementia", "Normal Cognition", "MCI", "Dementia", "Normal Cognition", "MCI", "Dementia")
  )

  data |>
    semi_join(filter_data, by = c("from", "to")) |>
    filter(Wave == wave) |>
    ggplot(aes(x = to, y = from, fill = transition_prob)) +
    geom_tile() +
    geom_text(aes(label = transition_prob), size = 5) +
    scale_fill_viridis_c(alpha = 0.8)+
    labs(title = paste0("Transition Probabilities Across Cognitive States (", wave, "-", wave + 2, ")"),
         x = "To", y = "From", fill = "Prob") +
    theme_light() +
    theme(
      plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),
      axis.title = element_text(size = 10),
      axis.text = element_text(size = 8),
      panel.grid = element_blank(),
      panel.border = element_blank()
    )
}

backtrack_age <- function(data, base_year, current_year) {
  # Back calculates an individual's age by 2-year intervals
  # Arguments:
  #   - data: The dataset containing the age variable and a 'Wave' column indicating the year of the wave
  #   - current_year: The year for which the age is measured (e.g., 2022)
  # Returns:
  #   - A dataset where the age is subtracted by intervals of 2 for each respective year

  # Map Wave values (1-10) to their corresponding years
  wave_to_year <- function(wave) {
    base_year + (as.numeric(wave) - 1) * 2
  }

  data |>
    group_by(ID) |>
    mutate(Age = Age - (current_year - wave_to_year(Wave))) |>
    ungroup()
}

pivot_data <- function(data, cog_total = FALSE) {
  # Reshapes the input dataset by pivoting cardiovascular, depression, and cognitive data
  # into long format for analysis. Handles two scenarios: one for cognitive status and
  # another for total cognitive scores. Combines the pivoted data into a final dataset
  # for further analysis.
  # Arguments:
  #   - data: The input dataset containing cardiovascular, depression, and cognitive data.
  #   - cog_total: A logical flag indicating whether to pivot total cognitive scores (TRUE)
  #               or cognitive status (FALSE).
  # Returns:
  #   - A reshaped dataset in long format with combined cardiovascular, depression, and
  #     cognitive data, ready for analysis.

  # Pivot cardio dataset
  data_cardio <- data |>
    tidyr::pivot_longer(
      cols = starts_with("Cardio_risk"),
      names_to = "Cardio_Year",
      values_to = "Cardio_risk"
    ) |>
    select(!any_of(c(
      "Cardio_Year", 
      paste0("cogfunction", 2016:2022), 
      paste0("cog_score_", 2016:2022)
      )))

  # Pivot dementia dataset
  data_dep <- data |>
    tidyr::pivot_longer(
      cols = starts_with("Total_dep"),
      names_to = "Dep_Year",
      values_to = "Depression_score"
    ) |>
    select(ID, Depression_score)

  if(cog_total == FALSE) {

    # Pivot cognition dataset
    data_cog <- data |>
      tidyr::pivot_longer(
        cols = starts_with("cogfunction"),
        names_to = "Wave",
        values_to = "Status"
      ) |>
      select(!any_of(c(
        paste0("Cardio_risk_", 16:22), 
        paste0("Total_dep_20", 16:22))
        ))

  # Adding cardio data to cognition data
  combine_data <- data_cog |>
    mutate(Cardio_risk = data_cardio$Cardio_risk, .after = Education_tri) |>
    mutate(Depression = data_dep$Depression_score, .after = Cardio_risk)

  # Producing final dataset
  final_data <- combine_data |>
    mutate(
      Wave = stringr::str_replace(Wave, "cogfunction", "HRS_"),
      Wave = factor(
        Wave,
        levels = c("HRS_2016", "HRS_2018", "HRS_2020", "HRS_2022"),
        labels = c(1, 2, 3, 4)
      )) |>
    group_by(ID) |>
    mutate(
      Status = ifelse(cumany(Status == "Dementia"), "Dementia", Status),
      Status = factor(
        Status,
        levels = c("Normal Cognition", "MCI", "Dementia"),
        labels = c(1, 2, 3)
      )) |>
    ungroup() |>
    mutate(
      Gender = factor(Gender, levels = c(0, 1)),
      Education_tri = factor(Education_tri, levels = c(0, 1, 2)),
      Wave = as.numeric(Wave),
      Status = as.numeric(Status)
    )
  } else {

    # Pivot cog total data
    data_score <- data |>
      tidyr::pivot_longer(
        cols = starts_with("cog_score"),
        names_to = "Wave",
        values_to = "Score"
      ) |>
      mutate(
        Wave = stringr::str_replace(Wave, "cog_score_", ""),
        Wave = factor(
          Wave, 
          levels = c("2016", "2018", "2020", "2022"),
          labels = c(1, 2, 3, 4)),

        Gender = as.factor(Gender),
        Education_tri = as.factor(Education_tri)
      ) |>
      select(!any_of(c(
        paste0("Cardio_risk_", 16:22), 
        paste0("Total_dep_20", 16:22))
      ))

    # Combine cardio and cog total data
    final_data <- data_score |>
      mutate(Cardio_risk = data_cardio$Cardio_risk, .after = Education_tri) |>
      mutate(Depression = data_dep$Depression_score, .after = Cardio_risk)
    }

  return(final_data)
}

gentleman_test <- function(data, model) {
  # Compares observed and expected prevalence of cognitive states (Normal Cognition, MCI, Dementia)
  # over time. Reshapes the observed and expected data into a long format and generates a
  # faceted line plot to visualize the comparison.
  # Arguments:
  #   - data: A list containing observed and expected prevalence data.
  #   - model: A string representing the model name, used in the plot title.
  # Returns:
  #   - A ggplot object displaying the observed vs. expected prevalence of cognitive states
  #     over time.

  # Reshaping observed prevalence
  do1 <- as_tibble(row.names(data$Observed)) |>
    rename(time = value) |>
    mutate(time = as.numeric(time))

  do2 <- as_tibble(data$Observed) |> 
    mutate(type = "observed")

  do <- cbind(do1, do2) |>
    select(-Total) |>
    tidyr::gather(state, number, -c(time, type)) |>
    mutate(state = case_when(state == "State 1" ~ "Normal Cognition",
                             state == "State 2" ~ "MCI",
                             state == "State 3" ~ "Dementia"))

  # Reshaping expected prevalence
  de1 <- as_tibble(row.names(data$Expected)) |>
    rename(time = value) |>
    mutate(time = as.numeric(time))

  de2 <- as_tibble(data$Expected) |>
    mutate(type = "expected")

  de <- cbind(de1,de2) |>
    select(-Total) |>
    tidyr::gather(state, number, -c(time, type))

  # Turn into single data frame and plotting
  rbind(do, de) |>
    mutate(type = factor(type),
           state = factor(state, levels = c("Normal Cognition", "MCI", "Dementia")),
           time = round(time,3)) |>
    group_by(state) |>
    ggplot(aes(x = time, y = number, colour = type)) +
    geom_line(linewidth = 1) +
    labs(title = paste0("Observed vs Fitted Values: ", model), x = "Time", y = "") +
    facet_wrap(~ state) +
    theme_bw() +
    ggeasy::easy_remove_legend_title()
}

process_hr <- function(data, fix_upper = TRUE) {
  # Processes hazard ratio (HR) data by cleaning and reshaping it for visualization.
  # Converts covariate names into descriptive labels and ensures proper formatting
  # of transitions and confidence intervals. Optionally fixes upper confidence intervals
  # that exceed a specified threshold.
  # Arguments:
  #   - data: The input dataset containing hazard ratio data.
  #   - fix_upper: A logical flag indicating whether to fix upper confidence intervals
  #                that exceed a threshold (default: TRUE).
  # Returns:
  #   - A cleaned and reshaped dataset ready for hazard ratio visualization.

  data <- do.call(rbind.data.frame, data) |>
    tibble::rownames_to_column(var = "Covariate") |>
    tidyr::separate_wider_delim(cols = Covariate, 
                                delim = ".", 
                                names = c("Covariate", "Transition")) |>
    mutate(Covariate = case_when(
      Covariate == "Gender1" ~ "Female",
      Covariate == "Education_tri1" ~ "High School",
      Covariate == "Education_tri2" ~ "College",
      Covariate == "Cardio_risk" ~ "Cardiovascular Risk",
      Covariate == "Depression" ~ "Depression",
      Covariate == "Total_p" ~ "Total Procrastination",
      Covariate == "Depression:Total_p" ~ "Depression x Procrastination",
      TRUE ~ Covariate),

      Covariate = factor(
        Covariate, 
        levels = c("Female", "Age", "High School", "College", 
                   "Cardiovascular Risk", "Depression", 
                   "Total Procrastination", "Depression x Procrastination"))
      ) |>
    mutate(Transition = factor(
      Transition, 
      levels = c("Normal Cognition - MCI", 
                 "Normal Cognition - Dementia",
                 "MCI - Normal Cognition",
                 "MCI - Dementia")))

  if(fix_upper == TRUE){
    data <- data |>
      mutate(U = case_when(
        U >= 5 & Transition == "Normal Cognition - Dementia" ~ HR,
        # U >= 5 & HR < 1 ~ 1,
        TRUE ~ U
      ))
  }

  return(data)
}

calculate_p_value <- function(data) {
  # Calculates p-values for hazard ratios (HR) based on log-transformed HR values and
  # their confidence intervals. Adds the calculated p-values to the input dataset.
  # Arguments:
  #   - data: The input dataset containing hazard ratios and confidence intervals.
  # Returns:
  #   - The input dataset with an additional column for p-values

  log_HR <- log(data$HR)
  log_lower_CI <- log(data$L)
  log_upper_CI <- log(data$U)

  SE <- (log_upper_CI - log_lower_CI) / (2 * 1.96)
  Z <- log_HR / SE
  p_value <- 2 * (1 - pnorm(abs(Z)))

  data$pval <- p_value |> round(digits = 3)

  return(data)
}

plot_hr <- function(data) {
  # Generates a forest plot of hazard ratios (HR) for cognitive transitions, including
  # confidence intervals and significance stars. Customizes the plot based on covariate
  # labels and ensures proper formatting for visualization.
  # Arguments:
  #   - data: The input dataset containing hazard ratios, confidence intervals, and p-values.
  # Returns:
  #   - A ggplot object displaying hazard ratios for cognitive transitions, faceted by covariate.

  # Making a plot customization
  labeller <- c(
    Female = "Being female vs. Male",
    `High School` = "Having a highshool education vs. No education",
    College = "Having a further education vs. No education",
    `Cardiovascular Risk` = "Total number of cardiovascular risk factors",
    Depression = "Depression",
    Total_p = "Procrastination",
    `Depression x Procrastination` = "Depression x Procrastination"
  )

  # Add a column for significance stars
  data <- data |>
    mutate(
      stars = case_when(
        pval < 0.001 ~ "***",
        pval < 0.01 ~ "**",
        pval < 0.05 ~ "*",
        pval < 0.10 ~ "+",
        TRUE ~ "" 
      ))

  # Plotting  
  haz_plot <- data |>
    ggplot(aes(x = HR, y = forcats::fct_rev(Transition), colour = HR > 1)) +
    geom_point(size = 4) +
    geom_linerange(aes(xmin = L, xmax = U), linewidth = 1.5, alpha = 0.5) +
    geom_vline(xintercept = 1, colour = "gray75", linewidth = 0.75, linetype = "dashed") +
    geom_text(aes(label = stars), vjust = -0.5, hjust = 0.5, size = 5, color = "black") +
    labs(title = "Hazard Ratio for dementia transitions",
         x = "Hazard Ratio (95% Confidence Interval)", 
         y = "") +
    scale_color_manual(values = c("#E69F00", "#56B4E9"), guide = "none") +
    guides(colour = "none") +
    theme_bw() +
    theme(
      plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
      axis.title.x = element_text(size = 12),
      strip.text = element_text(size = 10)
      )

  if(nrow(data) == 32){
  # Centering the bottom facet because I am pedantic
    haz_plot <- haz_plot + ggh4x::facet_manual(
    ~ Covariate,
    scales = "free_x",
    labeller = labeller(Covariate = labeller),
    design = c(
    "
    AABBCC
    DDEEFF
    #GGHH#
    "
    ))
  } else {
    # Centering the bottom facet because I am pedantic
    haz_plot <- haz_plot + ggh4x::facet_manual(
      ~ Covariate,
      scales = "free_x",
      labeller = labeller(Covariate = labeller),
      design = c(
      "
      AABBCC
      DDEEFF
      ##GG##
      "
      ))
  }

  return(haz_plot)
}
```

## Reading in data

```{r data}
set.seed(34521)

data <- read.csv(here::here("analysis/data/data.csv")) 
```

# Markov Modelling

In this section, we apply Markov modelling to analyze the transitions between cognitive states (Normal Cognition, Mild Cognitive Impairment [MCI], and Dementia) using longitudinal data from the Health and Retirement Study (HRS) from 2016 to 2022. By modeling these transitions, we aim to understand the probabilities of moving between different cognitive states over time.

## Data preparation

The first step is to prepare the transition dataset. We start by transforming the data to create a record of transitions from one cognitive state to another between consecutive waves.

Once the transitions are identified, we calculate the probabilities of each state-to-state transition. These probabilities are computed by counting the occurrences of each transition and dividing by the total number of transitions.

```{r transition-probability}
tran_prob <- data |>
  extract_years(years = seq(2016, 2022, by = 2)) |>
  create_transitions(absorbing = TRUE) |>
  calculate_probabilties()

tran_prob |>
  mutate(prop = round(prop, digits = 3)) |>
  head(n = 9) |>
  gt::gt() |>
  gtExtras::gt_theme_538() |>
  gt::cols_align(align = "center") |>
  gt::tab_header(title = "Overall Transition Probabilties (2016 - 2022)") |>
  gt::tab_options(table.width = gt::pct(100)) |>
  gt::opt_align_table_header(align = "center")
```

## Creating transition matrix

The probability distribution of transitions from one state to another can be represented into a transition matrix $P = (p_{ij})_{i,j}$ where each element of position $(i, j)$ represents the transition probability $p_{ij}$.

$$
P = \begin{bmatrix} 
p_{11} & p_{12} & p_{13} \\
p_{21} & p_{22} & p_{23} \\
p_{31} & p_{32} & p_{33} \\
\end{bmatrix}
$$

```{r transition-matrix}
tran_matrix <- tran_prob |>
  transition_matrix() |>
  normalise() |>
  round(digits = 3)

tran_matrix
```

This code creates a $3 \times 3$ matrix (for the three states: Normal Cognition, MCI, and Dementia) and populates it with the transition probabilities.

## Visualising the matrix

To make the transition matrix more interpretable we visualize the probabilities using a heat map (@fig-transition-matrix)

```{r transition-matrix-plot}
#| fig-width: 8
#| label: fig-transition-matrix
#| fig-cap: Heat map of transition probabilities (2016 - 2022)

plot_matrix(data = t(tran_matrix), year = 2016)
```

We can also create and plot transition matrices for each wave separately allowing us to see how these transition probabilities change across multiple waves of data collection

```{r transition-matrix-longitudinal}
tran_prob_long <- data |>
  extract_years(years = seq(2016, 2022, by = 2)) |>
  create_transitions(absorbing = TRUE) |>
  group_by(Wave) |>
  calculate_probabilties() |>
  group_map(~ transition_matrix(.x, longitudinal = TRUE), .keep = TRUE) |>
  lapply(function(mat) {
    mat[c("Normal Cognition", "MCI", "Dementia"), c("Normal Cognition", "MCI", "Dementia")] |>
        normalise()
    }) |>
  purrr::map2_df(.y = c(seq(2016, 2020, by = 2)), .f = reshape_matrix) |>
  mutate(
    transition_prob = round(transition_prob, digits = 3),
    from = factor(from, levels = c("Normal Cognition", "MCI", "Dementia"))
  )

DT::datatable(
  tran_prob_long,
  filter = "top",
  colnames = c("Probability" = "transition_prob"),
  options = list(pageLength = 6, autoWidth = TRUE)
)
```

```{r transition-matrix-longitudinal-plot, fig.width=8, fig.show="animate", ffmpeg.format="gif", dev="png"}
for(wave in c(seq(2016, 2020, by = 2))){
  print(plot_transition_wave(data = tran_prob_long, wave = wave))
}
```

# Multi-State Markov Model

Now that we have an idea of the disease progression across time we are interested in the effect of different covariates on these transitions. For this, we can fit a multi-state Markov model. Multi-state models (MSMs) provide a comprehensive view of the disease process and make risk-factors analysis and long-term predictions more easily. The conditional distribution of the status of an individual participant at an arbitrary examination given their status at previous examinations was assumed to have the **Markov property**. That is the distribution of the forthcoming state $X_{n+1}$ depends only on the current state $X_n$ and doesn't depend on the previous ones $X_{n-1}, X_{n-2}, \dots, X_1$.

$$Pr(X_{n+1} = x_{n+1} \; \vert \; X_1 = x_1, X_2 = x_2, \dots, X_n = x_n) = Pr(X_{n+1} = x_{n+1} \; \vert \; X_n = x_n)$$

## Data preperation

For our model we are going to look at several different covariates:

- **Gender**: $0 = \text{Male;} \; 1 = \text{Female}$ 
- **Age**: $\text{Range} \; = 50 - 91$ 
- **Education**: $0 = \text{No Degree;} \; 1 = \text{High School Degree;} \; 2 = \text{Further Education}$ 
- **Number of cardiovascular risk factors**: $\text{Range} \; = 0 - 4$
- **Procrastination**: $\text{Range} \; = 4 - 60$

```{r markov-prep}
cols <- c(
  "ID", "Gender", "Age", "Education_tri", 
  paste0("Cardio_risk_", seq(16, 22, by = 2)), 
  paste0("Total_dep_20", seq(16, 22, by = 2)), "Total_p")

markov_data <- data |>
  extract_years(seq(2016, 2022, by = 2)) |>
  na.omit() |>
  inner_join(data[, cols], by = "ID") |>
  pivot_data() |>
  backtrack_age(base_year = 2016, current_year = 2022) |>
  group_by(ID) |>
  mutate(
    Age = min(Age), # Baseline age
    Cardio_risk = Cardio_risk[Wave == 1],
    Depression = Depression[Wave == 1]
    ) |>
  ungroup() |>
  filter(Age >= 50) # Only looking at older adults

DT::datatable(
  markov_data,
  filter = "top",
  options = list(pageLength = 6, autoWidth = TRUE)
)
```

Let's look at a state table to visualize the amount of transitions over time

```{r state-table}
markov_data |>
  select(ID, Wave, Status) |>
  mutate(Status = case_when(
    Status == 1 ~ "Normal Cognition", 
    Status == 2 ~ "MCI", 
    Status == 3 ~ "Dementia")) |>
  msm::statetable.msm(state = Status, subject = ID)
```

## Modelling

### Defining a q-matrix

The Q matrix (also known as the transition intensity matrix or generator matrix) defines the rates at which transitions occur between states in the model.

It is a square matrix where:

- Each row corresponds to a **from** state in the model.
- Each column corresponds to a **to** state in the model.
- The off-diagonal elements $q_{ij}$ represent the transition rate from state $i$ to state $j$.

For a model with $n$ states, the Q matrix has the following structure

$$
Q = \begin{bmatrix} 
q_{11} & q_{12} & \cdots & q_{1n} \\
q_{21} & q_{22} & \cdots & q_{2n} \\
\vdots & \vdots & \ddots &\vdots  \\
q_{n1} & q_{n2} & \cdots & q_{nn}
\end{bmatrix}
$$

```{r q-matrix}
# Creating transition matrix
# Define possible transitions
states <- c("Normal Cognition", "MCI", "Dementia")

# Theoretical transition probabilities
q_matrix <- rbind(
  c(0.75, 0.20, 0.05),  # From Normal, can transition to MCI or Dementia
  c(0.25, 0.50, 0.25),  # From MCI, can transition to Normal or Dementia
  c(0.00, 0.00, 1.00)         # From Dementia, no transitions (absorbing state)
)

rownames(q_matrix) <- states
colnames(q_matrix) <- states

q_matrix
```

## Fitting

We will fit three different multi-state models

1. A model with no covariates
2. A multivariate model with covariates acting on all transitions
3. A multivariate model with covariates (and interactions) acting on all transitions
4. A multivariate model with covariates acting on only NC <-> MCI & MCI - Dementia
5. A multivariate model with covariates (and interactions) acting on only NC <-> MCI & MCI - Dementia

```{r model-fitting}
#| warning: false

# Fitting model with no covariates
model_1 <- msm::msm(
  formula = Status ~ Wave,
  subject = ID,
  data = markov_data,
  qmatrix = q_matrix,
  obstype = 1,
  method = "BFGS", 
  gen.inits = TRUE,
  control = list(
    fnscale = 5000, 
    maxit = 2000 # To reach convergence
  ))

# Multivariate model: covariates acting on all transitions
model_2 <- msm::msm(
  formula = Status ~ Wave,
  subject = ID,
  data = markov_data,
  qmatrix = q_matrix,
  covariates = ~ Gender + Age + Education_tri + Cardio_risk + Depression + Total_p,
  obstype = 1,
  gen.inits = TRUE,
  method = "BFGS",
  control = list(
    fnscale = 6000, 
    maxit = 2000 # To reach convergence
  ))

# Multivariate model: covariates acting on all transitions (with interaction)
model_3 <- msm::msm(
  formula = Status ~ Wave,
  subject = ID,
  data = markov_data,
  qmatrix = q_matrix,
  covariates = ~ Gender + Age + Education_tri + Cardio_risk + (Depression * Total_p),
  obstype = 1,
  gen.inits = TRUE,
  method = "BFGS",
  control = list(
    fnscale = 5000, 
    maxit = 2000 # To reach convergence
  ))

# Multivariate model: covariates acting on NC <-> MCI and MCI -> D
model_4 <- msm::msm(
  formula = Status ~ Wave,
  subject = ID,
  data = markov_data,
  qmatrix = q_matrix,
  covariates = list(
    "1-2" = ~ Gender + Age + Education_tri + Cardio_risk + Depression + Total_p,
    "2-1" = ~ Gender + Age + Education_tri + Cardio_risk + Depression + Total_p,
    "2-3" = ~ Gender + Age + Education_tri + Cardio_risk + Depression + Total_p
  ),
  obstype = 1,
  gen.inits = TRUE,
  method = "BFGS",
  control = list(
    fnscale = 5000, 
    maxit = 2000 # To reach convergence
  ))

# Multivariate model: covariates acting on NC <-> MCI and MCI -> D (with interaction)
model_5 <- msm::msm(
  formula = Status ~ Wave,
  subject = ID,
  data = markov_data,
  qmatrix = q_matrix,
  covariates = list(
    "1-2" = ~ Gender + Age + Education_tri + Cardio_risk + (Depression * Total_p),
    "2-1" = ~ Gender + Age + Education_tri + Cardio_risk + (Depression * Total_p),
    "2-3" = ~ Gender + Age + Education_tri + Cardio_risk + (Depression * Total_p)
  ),
  obstype = 1,
  gen.inits = TRUE,
  method = "BFGS",
  control = list(
    fnscale = 5000, 
    maxit = 2000 # To reach convergence
  ))

```

### Model Checking

After fitting each of our models we can extract some goodness of fit criteria

```{r model-goodness}
log_l <- c(logLik(model_1), 
           logLik(model_2), 
           logLik(model_3), 
           logLik(model_4), 
           logLik(model_5))

aic <- AIC(model_1, model_2, model_3, model_4, model_5)

# Very small difference between model 2 and 3 (makes sense)
cbind(log_l, aic) |>
  tibble::rownames_to_column("Model") |>
  mutate(Model = stringr::str_remove(Model, "model_"))

# Likelihood ratio test
lmtest::lrtest(model_1, model_2, model_3, model_4, model_5)
```

We can also look at a measure of how well the model fits the data using a test proposed by [@gentleman1994]. The function plots the observed numbers of individuals occupying a state at a series of times against forecasts from the fitted model, for each state (@fig-gentleman).

```{r gentleman_test}
#| fig-width: 12
#| fig-height: 12
#| label: fig-gentleman
#| fig-cap: Fitted values versus observed values

# Plotting fitted vs. observed values
g1 <- gentleman_test(data = msm::prevalence.msm(model_1), 
                     model = "Model 1")
g2 <- gentleman_test(data = msm::prevalence.msm(model_2), 
                     model = "Model 2")
g3 <- gentleman_test(data = msm::prevalence.msm(model_3), 
                     model = "Model 3")
g4 <- gentleman_test(data = msm::prevalence.msm(model_4), 
                     model = "Model 4")
g5 <- gentleman_test(data = msm::prevalence.msm(model_5), 
                     model = "Model 5")

ggpubr::ggarrange(g1, g2, g3, g4, g5, 
                  common.legend = TRUE, 
                  nrow = 5,
                  heights = c(2, 2, 2, 2, 2), 
                  legend = "bottom")
```

### Hazard ratios

We now present hazard ratios for each of our fitted models. These ratios represent how the instantaneous risk of making a particular transition is modified by the covariate.

::: {.panel-tabset}

## Model 2

@fig-hazard-ratio-1 presents hazard ratios from model 2 (our multivariate model with covariates acting on all transitions) for each covariate.

```{r}
#| fig-width: 12
#| fig-height: 12
#| label: fig-hazard-ratio-1
#| fig-cap: Ouputted hazard ratios from Markov model. Green bars represent a decreased risk of transition while red bars indicate an increased risk of transition.

# Plotting hazard ratios
model_2 |>
  msm::hazard.msm() |>
  process_hr() |>
  calculate_p_value() |>
  plot_hr()
```

## Model 3

@fig-hazard-ratio-2 presents hazard ratios from model 3 (our multivariate model with covariates **(and interactions)** acting on all transitions) for each covariate.

```{r}
#| fig-width: 12
#| fig-height: 12
#| label: fig-hazard-ratio-2
#| fig-cap: Ouputted hazard ratios from Markov model. Green bars represent a decreased risk of transition while red bars indicate an increased risk of transition.

# Plotting hazard ratios
model_3 |>
  msm::hazard.msm() |>
  process_hr() |>
  calculate_p_value() |>
  plot_hr()
```

## Model 4

@fig-hazard-ratio-3 presents hazard ratios for model 4 (our multivariate model with covariates acting on only NC <-> MCI & MCI -> Dementia) for each covariate

```{r}
#| fig-width: 12
#| fig-height: 12
#| label: fig-hazard-ratio-3
#| fig-cap: Ouputted hazard ratios from Markov model. Green bars represent a decreased risk of transition while red bars indicate an increased risk of transition.

# Plotting hazard ratios
model_4 |>
  msm::hazard.msm() |>
  process_hr() |>
  calculate_p_value() |>
  plot_hr()
```

## Model 5

@fig-hazard-ratio-4 presents hazard ratios for model 5 (our multivariate model with covariates **(and interactions)** acting on only NC <-> MCI & MCI -> Dementia) for each covariate

```{r}
#| fig-width: 12
#| fig-height: 12
#| label: fig-hazard-ratio-4
#| fig-cap: Ouputted hazard ratios from Markov model. Green bars represent a decreased risk of transition while red bars indicate an increased risk of transition.

# Plotting hazard ratios
model_5 |>
  msm::hazard.msm() |>
  process_hr() |>
  calculate_p_value() |>
  plot_hr()
```

:::