### 2025-03-28 analysis/data-final.qmd
Launching site

---
title: "Download final data"
format:
  html:
    code-tools: false
    toc: false
---

The cleaned data is available in three formats:

- CSV file for any program
- `.rds` file for R <span class="small">(load with `df <- readRDS("data.rds")`)</span>
- `.dta` file for Stata <span class="small">(load with `use "data.dta"`)</span>

## Data

<a class="btn btn-primary btn-sm" target="_blank" href="data/data.csv">{{< fa table >}} &ensp;<span class="font-monospace">data.csv</span></a> &emsp; <a class="btn btn-success btn-sm" target="_blank" href="data/data.rds">{{< fa brands r-project >}} &ensp;<span class="font-monospace">data.rds</span></a> &emsp; <a class="btn btn-warning btn-sm" target="_blank" href="data/data.dta">{{< fa table-cells >}} &ensp;<span class="font-monospace">data.dta</span></a>