### 2025-04-01 README.qmd
Update readme

---
format: gfm
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r setup, include=FALSE}
library(emoji)
```

##  `r emoji_glue(":potato::spiral_shell:")`: Note on "crisp clam" project name

Because project titles change all the time we used [{codename}](http://svmiller.com/codename/) to generate an [Ubuntu-style](https://wiki.ubuntu.com/DevelopmentCodeNames)
internal-to-us project name that won’t change.

```{r code-name, eval=FALSE}
library(codename)

codename_message()
#> code name generated by {codename} v.0.5.0. R version 4.4.2 (2024-10-31 ucrt).

codename(seed = "dementia", type = "ubuntu")
#> [1] "crisp clam"
```