### 2025-05-21 manuscript/manuscript.qmd
Update manuscript

---
title: "Procrastination and Dementia"
author: 
  - name: "Cormac Monaghan"
    url: "https://c-monaghan.github.io/"
    orcid: "https://orcid.org/0000-0001-9012-3060"
    affiliation:
      - "Hamilton Institute, Maynooth University, Ireland"
      - "Department of Psychology, Maynooth University, Ireland"
    email: cormacmonaghan@protonmail.com
    attributes:
      corresponding: true
  - name: "Rafael de Andrade Moral"
    url: "https://rafamoral.github.io/"
    orcid: "https://orcid.org/0000-0002-0875-3563"
    affiliation:
      - "Department of Mathematics and Statistics, Maynooth University, Ireland"
  - name: "Joanna McHugh Power"
    orcid: "https://orcid.org/0000-0002-7387-3107"
    affiliation: 
      - "Department of Psychology, Maynooth University, Ireland"
correspondence-prefix: "Correspondence concerning this article should be addressed to"
code-repo: "Access the code, data, and analysis at <https://github.com/C-Monaghan/crisp-clam>"
bibliography: references.bib
csl: csl/apa.csl
link-citations: true
---

::: {.content-visible when-format="html"}
# Introduction {.hide}
:::

Dementia is a syndrome characterized initially by the progressive and typically irreversible decline of cognitive function, leading to memory loss, impaired reasoning, and difficulties with activities of daily living [@Prince2013; @Sanz-Blasco2022]. It arises from various conditions, including Alzheimer's disease, vascular dementia, and Lewy body dementia [@Cao2020]. The global burden of dementia is substantial, with projections suggesting that the number of affected individuals will rise from 57.4 million in 2019 to 152.8 million by 2050 [@Nichols2022]. Given this projection, identifying and addressing modifiable risk factors is crucial to mitigate the growing prevalence difficulties with activities of daily living of dementia worldwide. In fact, targeting these factors could prevent up to 50% of dementia cases worldwide [@Livingston2024]. The *Lancet* Commission highlights 14 key modifiable risks, emphasizing that dementia risk can be reduced at any age. Nevertheless, age remains the strongest non-modifiable risk factor, accounting for the largest proportion of dementia cases.

Mild cognitive impairment (MCI) is a condition characterized by cognitive changes --- such as memory lapses or difficulty making decisions --- that exceed typical age-related decline [@Gauthier2006]. Prior studies suggest that 46% of people with MCI transition to a dementia state within three years and 80% within six years [@Cooper2015; @Shigemizu2020; @Tschanz2006]. However, progression is neither uniform nor inevitable. Heterogeneity in outcomes underscores the importance of distinguishing between two classes of predictors: causal risk factors (e.g. smoking, hypertension, or diabetes) which may accelerate dementia onset and are potential targets for intervention [@Livingston2024], and prodromal markers (e.g. memory loss, apathy, or struggles in daily living) that reflect early disease processes and may signal limited opportunities for prevention. While causal factors offer avenues for mitigating risk, prodromal features often emerge closer to irreversible pathological changes, highlighting the need for early detection and tailored interventions.

Apathy, defined as a significant loss of motivation, distinct from depression and cognitive impairment [@Fresnais2023; @Richard2012] is among the most prevalent neuropsychiatric symptoms in both MCI and dementia [@Donovan2014; @Fresnais2023; @Salem2023; @vanDalen2018]. Individuals with apathy exhibit reduced goal-directed behavior and a diminished desire to pursue rewards or pleasure [@Fresnais2023]. Importantly, apathy has been identified as a significant predictor for the transition from MCI to dementia [@Ruthirakuhan2019; @Salem2023; @vanDalen2018]. For instance, a meta-analysis by @vanDalen2018 found that apathy almost doubles the risk of transitioning to dementia. Additionally, levels of apathy measured using both the *"Neuropsychiatric Inventory Questionnaire\"* [@Kaufer2000] and *"Apathy Evaluation Scale\"* [@Marin1991] have shown to be correlated with higher levels of amyloid burden and neurofibrillary tangles in individuals with dementia, suggesting a potential connection to underlying neuropathology [@Donovan2014; @Marshall2013].

While apathy primarily reflects a lack of motivation, procrastination involves a delay in action despite an intention to complete such action [@Steel2007]. Although both behaviours indicate disruptions in goal-directed activity, they differ in their phenomenology: apathy is marked by diminished drive or emotional engagement, whereas procrastination typically involves intact intention but impaired follow-through. These differences suggest that procrastination may capture aspects of executive dysfunction and emotional regulation not fully accounted for by apathy alone.

Both behaviours have been linked with impaired executive function, particularly in goal-oriented behaviours, decision-making, and self-regulation [@Gustavson2015; @Kawagoe2017; @steel2010], domains which are frequently compromised in individuals with MCI and dementia [@Kirova2015; @Stopford2012]. While apathy has been well-established as a prodromal marker and risk factor for dementia progression [@Ruthirakuhan2019; @Salem2023; @vanDalen2018], emerging evidence also suggests a potential link between procrastination and early cognitive decline. For instance, @Friden2020 and @Zhang2019 highlight that chronic procrastination is associated with dysfunction in both the dorsolateral and ventromedial prefrontal cortices, regions of the brain known to not only show early degeneration in dementia [@Joseph2021], but also be linked to higher levels of apathy [@Fahed2021; @Zhang2019].

Furthermore, procrastination may exacerbate existing cognitive decline by reinforcing patterns of inaction and passivity. Individuals who chronically delay tasks may inadvertently engage in fewer cognitively stimulating activities, such as physical activity, problem-solving, decision-making, and goal-setting. Such activities are known to build cognitive resilience and reduce dementia risk [@Chowdhary2022; @Livingston2024]. By limiting engagement in such activities, procrastination may contribute to a self-reinforcing cycle of cognitive disengagement, potentially accelerating decline over time.

Given these behavioural and neuro-cognitive parallels with apathy, it is worth exploring whether chronic procrastination could serve as an early behavioural marker for cognitive impairment, or even a risk factor for dementia, especially in older adults. By examining procrastination alongside apathy, we may uncover a broader behavioural phenotype indicative of early neuropathological changes. Crucially, procrastination may identify individuals with motivational or executive dysfunction who do not yet meet the threshold for apathy, offering a potentially valuable and earlier behavioural marker for neurodegenerative risk.

Moreover, since age remains the strongest predictor of MCI and dementia, it is important to consider how behavioural predictors such as procrastination operate across the lifespan. Many established modifiable risk factors such as, hypertension, hearing loss, smoking, and social isolation, exert age-dependent effects, with certain factors carrying more weight at midlife than in late life [@livingston2020; @Livingston2024]. In light of this, we sought to explore whether procrastination also demonstrates an age-dependent risk profile. Understanding how procrastination interacts with age may help clarify its role in the aetiology of cognitive decline and identify windows of opportunity for targeted intervention.

Although current research on procrastination in relation to dementia is non-existent, this possible association warrants exploration. Identifying procrastination as a potential risk factor could expand the scope of early interventions aimed at preventing or slowing the progression of dementia. As such, the purpose of this study was to (a) assess group differences in procrastination levels across the normal cognition, MCI, and dementia groups, and b) test the hypotheses that higher levels of procrastination would be associated with an increased probability of transitioning from normal cognitive function or MCI to dementia.

# Method

## Data and study population {#data-and-study-population .unnumbered}

Analyses were conducted using a secondary data source; a multi-wave prospective cohort study called the Health and Retirement Study (HRS; [@Fisher2018], which tracks the health, economic, and social well-being of over $18,000$ American adults primarily aged $>50$. The HRS is managed by the Institute for Social Research at the University of Michigan, with data collected every two years. Initial data collection of a participant is conducted through a face-to-face interview, with follow-up biennial interviews conducted either by phone or face-to-face. The average re-interview response rate ranges from $68.8\%$ to $92.3\%$ [@HRS2017]. At the time of writing, fifteen years of HRS data are currently archived.

For this study, we focused on four biennial waves of HRS data (from 2016 to 2022). Specifically our study sample consisted of respondents who participated in an experimental module assessing procrastination during the 2020 wave. These experimental modules, administered at the end of the core HRS interview, consist of concise questionnaires designed to explore new topics or supplement existing core survey data. Each respondent receives only one experimental module, with sample sizes for each module constituting approximately 10% of the core sample. As a result, while the core HRS sample includes approximately $18,000$ respondents, our initial sample of interest consisted of $1,368$ respondents. We excluded respondents with missing cognitive assessment data for any wave $(n = 445)$, missing procrastination data $(n = 4)$, and those under 50 years of age  $(n = 16)$. This resulted in a final analytic sample of $903$ respondents.

## Measures

### Outcome: Cognitive Function and Cognitive Category

Cognitive function in the HRS is assessed using a series of tests adapted from the Telephone Interview for Cognitive Status (TICS; [@Brandt1988; @Fong2009]. These tests include an immediate and delayed $10$-noun free recall test, a serial seven subtraction test, and a backward count from $20$ test. Based on these assessments, @Crimmins2011 developed both a $27$-point cognitive scale and validated cut-off points to assess and classify cognitive status. Using these points, respondents who scored $12 - 27$ were classified as having normal cognition, $7 - 11$ as having MCI, and $0 - 6$ as having dementia.

### Predictor: Procrastination

Procrastination was measured using the Pure Procrastination Scale [@steel2010], which consists of 12 items rated on a Likert scale ranging from $1$ (strongly disagree) to $5$ (strongly agree). The total procrastination score ranges from $12$ to $60$, with higher scores indicating greater procrastination. The Pure Procrastination Scale conducted in 2020 (wave 3) had a Cronbach's $\alpha$ score of $0.91$ in this sample, indicating high internal consistency. An example of a question from the scale includes "I delay making decisions until it's too late".

### Confounders

To account for potential confounding, we controlled for variables with established associations with cognitive function and procrastination. These included baseline measures conducted in 2016 (wave 1) of age, sex, educational attainment, number of cardiovascular risk factors (history of hypertension, stroke, or heart disease, and a classification on "overweight\" on a body mass index (BMI) scale $(\text{BMI} \ge 30)$), and depressive symptoms [@Abner2012; @Freedman2024; @monaghan2024a; @Yu2013]. Educational attainment was classified into three categories: no formal education, GED (General Educational Development)/high school diploma, and college/further education. Depressive symptoms were measured using an eight-item version of the Center for Epidemiological Studies Depression (CES-D) scale [@Briggs2018], with scores ranging from 0 to 8, demonstrating good internal consistency $(\alpha = 0.82)$.

# References