### 2025-06-24 analysis/discrete-markov.qmd
Update

---
title: "Discrete Time Markov Model"
---

# Setting up

```{r}
#| label: setup
#| warning: false
#| message: false
#| output: false
#| code-fold: true
#| code-summary: "Check out my code"

rm(list = ls())

# Packages ---------------------------------------------------------------------
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)

# Functions --------------------------------------------------------------------
files <- list.files(here::here("R/"), full.names = TRUE)

purrr::walk(files, source)

# Theme ------------------------------------------------------------------------
colour <- "#212427"

theme_set(
  theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, size = 14, colour = colour, face = "bold"),
      plot.subtitle = element_text(hjust = 0.5, size = 12, colour = colour),
      axis.title = element_text(size = 10, colour = colour, face = "bold"),
      strip.text = element_text(size = 10, colour = colour, face = "bold"),
      legend.title = element_text(hjust = 0.5, colour = colour, face = "bold"),
      ))

# Data -------------------------------------------------------------------------
data <- read.csv(here::here("analysis/data/data.csv"))
```

# Data Preparation

Before preparing the data, we want to create a vector of covariates that we are interested in using in our analysis

```{r covariates}
cols <- c(
  "ID", "Gender", "Age", "Education_tri",
  paste0("Total_dep_", seq(2016, 2022, by = 2)),
  "Total_p")
```

Following this, we can now use the `extract_years()` function to extract the cognitive function data for the years 2016, 2018, and 2020.

- We will also impute some missing values `impute = TRUE` using logical reasoning (if a respondent has an NA value in 2018, but has a classification of "normal cognition" in 2020, then the missing 2018 value becomes "normal cognition").
- We will also treat dementia as an absorbing state `absorbing = TRUE`.

```{r extract-years}
data_stack <- data |> 
  extract_years(years = seq(2016, 2022, by = 2), impute = TRUE, absorbing = TRUE) |>
  na.omit()

head(data_stack)
```

## Creating a stacked dataset

For each respondent we will now add in their relevant covariate data. Following this, we transform the data to long format and convert categorical variables to factors using the `pivot_and_factorise()` function. Additionally, we will fix the `Age` column to properly represent the age of the respondent at each time point

```{r long_data}
data_stack <- data_stack |>
  inner_join(data[, cols], by = "ID") |>
  # Removing baseline age below 60
  filter(Age - (2022 - 2016) >= 60) |>
  pivot_and_factorise(time_vary = TRUE) |>
  group_by(ID) |>
  # ^ Make age time varying
  mutate(Age = Age - (2022 - as.numeric(as.character(wave))))

head(data_stack)
```

Finally, we will create a new variable `status_prev` that notes the cognitive status of the respondent in the previous wave $(t - 1)$. This will be done by using the `lag` function from the `dplyr` package.

```{r stack-data}
data_stack <- data_stack |>
  mutate(status_prev = lag(status), .after = status) |>
  filter(!is.na(Total_p)) |>
  ungroup() |> 
  filter(wave != 2016)

head(data_stack)
```

# Transition frequencies

We will now calculate the transition frequencies between cognitive states for each time period. We will use the `create_transition_table()` function to create a transition table for each time period. This will be done with the help of the `map()` function from the `purrr` package. Finally, to combine all the transition tables into one dataset, we will use the `create_transition_dataset()` function.

```{r transition-tables}
# Creating a table dataset
table_data <- data |>
  extract_years(seq(2016, 2022, by = 2)) |>
  rename_with(~ gsub("cogfunction", "HRS_", .)) |>
  mutate(
    across(c(HRS_2016:HRS_2022), ~ factor(.x, levels = c("Normal Cognition", "MCI", "Dementia")))
  )

# Creating transition frequencies ---------------------------------------------
## These are the time periods we are interested in
time_periods <- list(
  c("2016", "2018"),
  c("2018", "2020"), 
  c("2020", "2022")
)

# Applying function
transition_results <- purrr::map(time_periods, ~ create_transition_table(.x[1], .x[2]))
names(transition_results) <- purrr::map_chr(time_periods, ~ paste(.x[2], .x[1], sep = "-"))

### Creating one dataset
transition_frequencies <- time_periods |>
  create_transition_dataset(transition_results = transition_results)

head(transition_frequencies)
```

## Observed transition matrix

The probability distribution of transitions from one state to another can be represented into a transition matrix $P = (p_{ij})_{i,j}$ where each element of position $(i, j)$ represents the transition probability $p_{ij}$.

In order to create this matrix we will use both the `create_transitions()` and `observed_transition_matrix()` functions.

```{r observed-matrix}
## Creating observed transition matrix -----------------------------------------
transition_matrix_observed <- data |>
  extract_years(seq(2016, 2022, by = 2)) |>
  create_transitions() |>
  observed_transition_matrix()

transition_matrix_observed
```

## Visualisation

Let's visualize both the transition frequencies (@fig-frequencies) and matrix (@fig-matrix).

```{r}
#| code-fold: true
#| code-summary: "Check out my code"
#| fig-width: 12
#| fig-height: 10
#| label: fig-frequencies
#| fig-cap: "Transition frequencies between cognitive states for each time period."

fig_1 <- transition_frequencies |> tran_stack_graph()
fig_2 <- transition_frequencies |> tran_heat_map()

fig_1 / fig_2
```

```{r}
#| code-fold: true
#| code-summary: "Check out my code"
#| fig-width: 12
#| label: fig-matrix
#| fig-cap: "Observed transition matrix between cognitive states for the years 2016 - 2022."

fig_3 <- transition_matrix_observed |>
  reshape_matrix() |>
  plot_transition_matrix()

fig_3
```

# Modelling

## Markov Process Fundamentals

Discrete-time Markov models belong to a class of stochastic processes that satisfy the **Markov property**, which can be formally expressed as:

$$
P(X_{t+1} = j \vert X_t = i, X_{t-1} = i_{t-1}, \dots X_0 = i_0) = P(X_{t+1} = j \vert X_t = i)
$$ {#eq-markov}

This property establishes that the future state $X_{t+1}$ depends only on the current state $X_t$, not on the entire history of states.

### Transition Probability Matrix

For our three-state system (Normal Cognition, Mild Cognitive Impairment [MCI], Dementia), the transition matrix $P$ captures all possible transition probabilities:

$$
P = \begin{bmatrix} 
p_{11} & p_{12} & p_{13} \\
p_{21} & p_{22} & p_{23} \\
p_{31} & p_{32} & p_{33} \\
\end{bmatrix}
$$ {#eq-t-matrix}

where:

- $p_{ij} = P(X_{t+1} = j \vert X_t = i)$ represents the probability of transitioning from state $i$ to state $j$.
- Each row sums to 1, $\sum^3_{j=1} p_{ij} = 1 \quad  \forall_i \in \{1, 2, 3\}$

### Multinomial Logistic Regression Formulation

We model the transition probabilities using **multinomial logistic regression**, where the log-odds of each transition relative to a reference state are linear functions of covariates.

For a system with $K$ states (using state $K$ as reference), we have:

$$
log \left( \frac{P(Y = j \vert x)}{P(Y = k \vert x)} \right) = \beta_{j0} + \beta_j^Tx \qquad \text{for } j = 1, \dots K-1
$$ {#eq-multinomial}

For non-reference states $j = 1, \dots, K - 1$

$$
P(Y = j \vert x) = \frac{e^{\beta_{0j} + \beta_j^Tx}}{1 + \sum^{k - 1}_{k = 1} e^{\beta_{0k} + \beta_k^Tx}}
$$ {#eq-non-reference}

For the reference state $K$:

$$
P(Y = k \vert x) = \frac{1}{1 + \sum^{k - 1}_{k = 1} e^{\beta_{0k} + \beta_j^Tx }}
$$ {#eq-reference}

### Time-Homogeneous Approach

Our implementation assumes time-homogeneous transitions, but the framework can be extended to time-varying probabilities:

$$
P^{(t)} = \begin{bmatrix} 
p_{11}^{(t)} & p_{12}^{(t)} & \cdots & p_{1n}^{(t)} \\
p_{21}^{(t)} & p_{22}^{(t)} & \cdots & p_{2n}^{(t)} \\
\vdots & \vdots & \ddots & \vdots \\
p_{n1}^{(t)} & \cdots & \cdots & p_{nn}^{(t)} \\
\end{bmatrix}
$$ {#eq-non-stationary}

Our primary analysis assumes the transition matrix remains fixed:

$$
P^{(t)} \equiv P \qquad \forall \; t
$$ {#eq-assumption}

### Absorbing State Specification

We model Dementia as an absorbing state:

$$
p_{3j} = \begin{cases} 
1 \qquad \text{if } j = 3 \\
0 \qquad \text{otherwise} \end{cases}
$$ {#eq-dementia}

yielding the constrained transition matrix:

$$
P = \begin{bmatrix} 
p_{11} & p_{12} & p_{13} \\
p_{21} & p_{22} & p_{23} \\
0      & 0      & 1      \\
\end{bmatrix}
$$ {#eq-t-matrix-constrained}

## Stationary model

We estimate three progressively complex stationary models using `nnet::multinom()`

```{r model-fitting}
#| code-fold: true
#| code-summary: "Check out my models"
#| 
# Baseline model (only gender; reference: NC)
fit_1a <- nnet::multinom(
  status ~ Gender + Age + Education_tri + Depression, 
  family = multinomial, 
  data = data_stack, trace = FALSE)

# Baseline model (only gender; reference: MCI)
fit_1b <- nnet::multinom(
  status ~ Gender + Age + Education_tri + Depression, 
  family = multinomial, 
  data = data_stack |> mutate(status = relevel(status, ref = 2)), 
  trace = FALSE)

# With procrastination (reference: NC)
fit_2a <- nnet::multinom(
  status ~ Gender + Age + Education_tri + Depression + Total_p, 
  family = multinomial, 
  data = data_stack, trace = FALSE)

# With procrastination (reference: MCI)
fit_2b <- nnet::multinom(
  status ~ Gender + Age + Education_tri + Depression + Total_p, 
  family = multinomial, 
  data = data_stack |> mutate(status = relevel(status, ref = 2)), 
  trace = FALSE)

# Full model with previous state (reference NC)
fit_3a <- nnet::multinom(
  status ~ Gender + Age + Education_tri + Depression + Total_p + status_prev, 
  family = multinomial, 
  data = data_stack, trace = FALSE)

# Full model with previous state (reference MCI)
fit_3b <- nnet::multinom(
  status ~ Gender + Age + Education_tri + Depression + Total_p + status_prev, 
  family = multinomial, 
  data = data_stack |> mutate(status = relevel(status, ref = 2)), 
  trace = FALSE)

# Full model with interaction (reference NC)
fit_4a <- nnet::multinom(
  status ~ Gender + Age + Education_tri + Depression + Total_p + (Total_p * Age) + status_prev, 
  family = multinomial, 
  data = data_stack, trace = FALSE)

# Full model with interaction (reference MCI)
fit_4b <- nnet::multinom(
  status ~ Gender + Age + Education_tri + Depression + Total_p + (Total_p * Age) + status_prev, 
  family = multinomial, 
  data = data_stack |> mutate(status = relevel(status, ref = 2)), 
  trace = FALSE)
```

### Model comparison

We evaluate model improvement (@tbl-stationary-comparison) using likelihood ratio tests:

$$
D = -2 \times \ell_{\text{reduced}} - \ell_{\text{full}}
$$ {#eq-liklihood}

```{r}
#| label: tbl-stationary-comparison
#| tbl-cap: "Likelihood ratio test for stationary models"
#| tbl-cap-location: bottom

anova(fit_1a, fit_2a, fit_3a, fit_4a) |>
  mutate(
    Model = c("Baseline", 
              "Procrastination", 
              "Procrastination + Previous Status",
              "Interaction Model"),
    `Resid. Dev`= round(`Resid. Dev`, digits = 3),
    `LR stat.` = round(`LR stat.`, digits = 3),
    `Pr(Chi)` = round(`Pr(Chi)`, digits = 3)
    ) |>
  DT::datatable(
    options = list(
      pageLength = 4,
      dom = "t",
      ordering = FALSE,
      columnDefs = list(list(className = "dt-center", targets = "_all"))
    ),
    rownames = FALSE
  )
```

::: {.panel-tabset .nav-pills}
## Non-interaction model

### Estimated parameters

Below we present the estimated parameters for the stationary model (fit 3) (@tbl-stationary-output). We will transform the estimates to odds ratios and colour them based on significance.

```{r}
#| code-fold: true
#| code-summary: "Check out my code"
#| label: tbl-stationary-output
#| tbl-cap: "Estimated parameters for the stationary model"
#| tbl-cap-location: bottom

stationary_results <- rbind(tidy_output(fit_3a), tidy_output(fit_3b)) |>
  mutate(
    # Mutating to factors
    transition = factor(
      transition, 
      levels = c("NC - MCI", "MCI - NC", 
                 "NC - Dementia", "MCI - Dementia")),
    term = factor(
      term, 
      levels = c("Being female", "Age", "High school degree vs. No education",
                 "Further education vs. No education", "Depression",
                 "Procrastination (2020)", "Previous state: MCI", 
                 "Previous state: Dementia"))
    )

stationary_results |>
  mutate(across(c(estimate, conf.low, conf.high), ~ round(x = ., digits = 3))) |>
  DT::datatable(
    options = list(
      pageLength = 6,
      dom = "tip",
      ordering = FALSE,
      columnDefs = list(list(className = "dt-center", targets = "_all"))
    ),
    rownames = FALSE
  )
```

### Model visualisation

Let's visualize the estimated odds ratios (@fig-odds-ratio) and the predicted transition probabilities (@fig-model-predictions) for the stationary model .

```{r}
#| message: false
#| warning: false
#| code-fold: true
#| code-summary: "Check out my code"
#| fig-width: 12
#| label: fig-odds-ratio
#| fig-cap: "Estimated odds ratios for the stationary model."

fig_4a <- stationary_results |>
  filter(!term %in% c("Previous state: MCI", "Previous state: Dementia")) |>
  ggplot(aes(x = estimate, y = transition, colour = colour)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray50") +
  ggstance::geom_pointrangeh(
    aes(xmin = conf.low, xmax = conf.high),
    position = ggstance::position_dodgev(height = 0.5),
    size = 1.25,
    fatten = 3) +
  scale_colour_manual(values = c(
    "Positive" = "#0072B2", 
    "Negative" = "#E69F00", 
    "NS"       = "#B2BEB5")) +
  labs(title = "Odds ratios (stationary model)",
       x = "Odds Ratio", y = "Predictor") +
  guides(colour = "none") +
  facet_wrap(~ term, scales = "free_x") +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    strip.text = element_text(size = 10, face = "bold"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank())

fig_4a
```

```{r}
#| code-fold: true
#| code-summary: "Check out my code"
#| fig-width: 12
#| label: fig-model-predictions
#| fig-cap: "Predicted transition probabilities from the stationary model."
#| 
## Making a prediction dataset -------------------------------------------------
pred_data <- expand.grid(
  Gender = factor(0),
  Age = mean(data_stack$Age),
  Education_tri = factor(0),
  Depression = mean(data_stack$Depression, na.rm = TRUE),
  status_prev = levels(data_stack$status_prev),
  Total_p = seq(0, 60, length = 200))

## Plotting predictions
fig_4b <- pred_data |>
  modelr::add_predictions(model = fit_3a, var = "pred", type = "probs") |>
  tidy_predictions() |>
  plot_predictions_stationary(variable = Total_p, x_axis = "Total Procrastination")

fig_4b
```

### Estimated transition matrix

Finally, we can compare the observed and predicted transition matrices. We will use the estimated probabilities from the model to fill in the transition matrix.

```{r estimated-matrix}
# Get all unique states
names <- c("Normal Cognition", "MCI", "Dementia")
states <- sort(unique(data_stack$status))
n_states <- length(states)

# Create empty transition matrix
transition_matrix_estimated <- matrix(
  0, nrow = n_states, ncol = n_states,
  dimnames = list(paste("From", names), paste("To", names))
  )

# Getting estimated probabilities
estimated_probs <- expand.grid(
  Gender = factor(0),
  Age = mean(data_stack$Age),
  Education_tri = factor(0),
  Depression = mean(data_stack$Depression, na.rm = TRUE),
  Total_p = mean(data_stack$Total_p),
  status_prev = states) |>
  modelr::add_predictions(model = fit_3a, var = "pred", type = "probs")

# Filling in matrix
for(i in 1:n_states) {
  transition_matrix_estimated[i, ] <- estimated_probs$pred[i, ]
}

transition_matrix_estimated |> round(digits = 3)
```

Let's plot this and then compare with our observed matrix

```{r}
#| code-fold: true
#| code-summary: "Check out my code"
#| fig-width: 12
#| label: fig-matrix-comparison
#| fig-cap: "Comparison of observed and estimated transition matrices."

fig_5 <- transition_matrix_estimated |>
  reshape_matrix() |>
  plot_transition_matrix(observed = FALSE)

fig_3 + fig_5 + plot_layout(axis_titles = "collect", guides = "collect")
```

## Interaction model

### Estimated parameters

Below we present the estimated parameters for the interaction model (fit 4; @tbl-interaction-output). We will transform the estimates to odds ratios and colour them based on significance.

```{r}
#| code-fold: true
#| code-summary: "Check out my code"
#| label: tbl-interaction-output
#| tbl-cap: "Estimated parameters for the interaction model"
#| tbl-cap-location: bottom

interaction_results <- rbind(tidy_output(fit_4a), tidy_output(fit_4b)) |>
  mutate(
    # Mutating to factors
    transition = factor(
      transition, 
      levels = c("NC - MCI", "MCI - NC", 
                 "NC - Dementia", "MCI - Dementia")),
    term = factor(
      term, =
      levels = c("Being female", "Age", 
                 "High school degree vs. No education",
                 "Further education vs. No education", "Depression",
                 "Procrastination (2020)",  
                 "Age x Procrastination", "Previous state: MCI", 
                 "Previous state: Dementia")),
    )

interaction_results |>
  mutate(across(c(estimate, conf.low, conf.high), ~ round(x = ., digits = 4))) |>
  DT::datatable(
    options = list(
      pageLength = 10,
      dom = "tip",
      ordering = FALSE,
      columnDefs = list(list(className = "dt-center", targets = "_all"))
    ),
    rownames = FALSE
  )
```

### Model visualisation

Let’s visualize the estimated odds ratios (@fig-odds-ratio-interaction) and predicted transition probabilities (@fig-interaction-predictions) for our interaction model.

```{r}
#| message: false
#| warning: false
#| code-fold: true
#| code-summary: "Check out my code"
#| fig-width: 12
#| fig-height: 10
#| label: fig-odds-ratio-interaction
#| fig-cap: "Estimated odds ratios for the interaction model."

# Design matrix for faceting
design <- "
  AABBCC 
  DDEEFF
  ##GG##
"

interaction_results |>
  filter(!term %in% c("Previous state: MCI", "Previous state: Dementia")) |>
  ggplot(aes(x = estimate, y = transition, colour = colour)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray50") +
  ggstance::geom_pointrangeh(
    aes(xmin = conf.low, xmax = conf.high),
    position = ggstance::position_dodgev(height = 0.5),
    size = 1.25,
    fatten = 3) +
  scale_colour_manual(values = c(
    "Positive" = "#0072B2", 
    "Negative" = "#E69F00", 
    "NS"       = "#B2BEB5")) +
  labs(title = "Odds ratios (stationary model)",
       x = "Odds Ratio", y = "Predictor") +
  guides(colour = "none") +
  ggh4x::facet_manual(~ term, scales = "free_x", design = design) +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    strip.text = element_text(size = 10, face = "bold"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank())

```

```{r}
#| code-fold: true
#| code-summary: "Check out my code"
#| warning: false
#| fig-width: 12
#| fig-height: 10
#| label: fig-interaction-predictions
#| fig-cap: "Predicted transition probabilities from the interaction model."

# Making a new predicted dataset
pred_data <- expand.grid(
  Gender = factor(0),
  Age = seq(60, 97, length = 200),
  Education_tri = factor(0),
  Depression = mean(data_stack$Depression, na.rm = TRUE),
  status_prev = levels(data_stack$status_prev),
  Total_p = seq(0, 60, length = 200))

# Reveling original data
data_plot <- data_stack |>
  mutate(
    # Relevel
    status = case_when(
      status == 1 ~ "Normal Cognition",
      status == 2 ~ "MCI",
      status == 3 ~ "Dementia"),
    status_prev = case_when(
      status_prev == 1 ~ "Normal Cognition",
      status_prev == 2 ~ "MCI",
      status_prev == 3 ~ "Dementia")
    ) |>
  mutate(
    status = factor(
      status, levels = c("Normal Cognition", "MCI", "Dementia")),
    status_prev = factor(
      status_prev, levels = c("Normal Cognition", "MCI", "Dementia"))) |>
  filter(status_prev %in% c("Normal Cognition", "MCI") & 
         status %in%  c("Normal Cognition", "MCI"))

# Plotting interaction effect ----------------------------------------------
pred_data |>
  # Adding predictions
  modelr::add_predictions(model = fit_4a, var = "pred", type = "probs") |>
  tidy_predictions() |>
  filter(status_prev %in% c("Normal Cognition", "MCI") & 
         status %in%  c("Normal Cognition", "MCI")) |>
  ggplot(aes(x = Total_p, y = as.numeric(Age))) +
  # Adding heat mat with contour lines
  geom_raster(aes(fill = prob), alpha = 0.7) +
  geom_contour(aes(z = prob), colour = "black", size = 0.3) +
  # Adding orginal data points
  geom_point(
    data = data_plot, aes(x = Total_p, y = Age), size = 1.75, alpha = 0.3) +
  # Customising plot
  scale_x_continuous(
    breaks = seq(0, 60, by = 10),
    sec.axis = sec_axis(~ ., name = "Previous State",
                        breaks = NULL, labels = NULL)) +
  scale_y_continuous(
    breaks = seq(50, 100, by = 5),
    sec.axis = sec_axis(~ ., name = "Current State",
                        breaks = NULL, labels = NULL)) +
  scale_fill_viridis_c(option = "plasma") +
  labs(
    title = "Predicted transition probabilities from the interaction model",
    x = "Total Procrastination", y = "Age", 
    fill = expression(hat(p))) +
  facet_grid(status ~ status_prev) +
  theme_classic() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    strip.text = element_text(size = 10, face = "bold"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.title = element_text(size = 10, face = "bold"))

```

Finally, (@fig-interaction-age-predictions) shows the the predicted transition probabilities for different age cohorts.

```{r}
#| code-fold: true
#| code-summary: "Check out my code"
#| warning: false
#| fig-width: 12
#| fig-height: 8
#| label: fig-interaction-age-predictions
#| fig-cap: "Predicted transition probabilities for different age cohorts."

# Create predictions with age info
age_cohots <- expand.grid(
  Gender = factor(0),
  Age = c(62:97),
  Education_tri = factor(0),
  Depression = mean(data_stack$Depression, na.rm = TRUE),
  status_prev = levels(data_stack$status_prev),
  Total_p = seq(0, 60, length = 200)) |>
  modelr::add_predictions(model = fit_4a, var = "pred", type = "probs") |>
  tidy_predictions() |>
  filter(status_prev %in% c("Normal Cognition", "MCI") &
         status %in%  c("Normal Cognition", "MCI")) |>
  mutate(label = case_when(
    Age %in% c(70, 80, 90) ~ as.character(Age), 
    TRUE ~ "Other"
  ))

# Identifying right most point for each label point
label_data <- age_cohots |>
  filter(Age %in% c(70, 80, 90)) |>
  group_by(status_prev, Age) |>
  filter(Total_p == max(Total_p)) |>
  ungroup()

# Plotting ---------------------------------------------------------------------
ggplot(data = age_cohots, aes(x = Total_p, y = prob, 
                              group = interaction(Age, status_prev, status),
                              colour = label)) +
  geom_line(aes(linewidth = label)) + 
  ggrepel::geom_text_repel(
    data = label_data,
    aes(label = paste0("Age ", Age), colour = as.character(Age)),
    direction = "y", nudge_x = 4, hjust = 0,
    segment.size = 0.5, segment.colour = "grey50",
    size = 3, show.legend = FALSE) +
  scale_x_continuous(
    breaks = seq(0, 60, by = 10),
    sec.axis = sec_axis( ~., name = "Previous State", breaks = NULL, labels = NULL)) +
  scale_y_continuous(sec.axis = sec_axis( ~ ., name = "Current State", breaks = NULL, labels = NULL)) +
  scale_colour_manual(values = c(
    "70"    = "#E69F00", 
    "80"    = "#56B4E9", 
    "90"    = "#009E73",
    "Other" = "grey80")) +
  scale_linewidth_manual(values = c(
    "70"    = 1.5, 
    "80"    = 1.5, 
    "90"    = 1.5,
    "Other" = 0.5)) +
  labs(
    title = "Predicted transition probabilities for different age cohorts",
    x = "Procrastination", y = "Probability", colour = NULL) +
  facet_grid(status ~ status_prev) +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    strip.text = element_text(size = 10, face = "bold"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.title = element_text(size = 10, face = "bold"),
    legend.position = "none")
```

:::

## Non-stationary model

We can also estimate a non-stationary model (@eq-non-stationary) by incorporating time `(wave)` into our model. This model allows for time-varying transition probabilities $p^{(t)}_{ij}$.

We estimate three progressively complex non-stationary models:

- Additive time effects `(fit_4)`: Baseline covariates + wave
- State-specific time effects `(fit_5)`: Interaction between previous state and wave
- Full time interactions `(fit_6)`: All covariates interacting with wave

```{r non-stationary}
#| code-fold: true
#| code-summary: "Check out my models"

## Small processing to fix wave column
data_stack <- data_stack |>
  mutate(wave = case_when(
    wave == "2016" ~ 1,
    wave == "2018" ~ 2,
    wave == "2020" ~ 3,
    wave == "2022" ~ 4
  ))

# Model 5: Additive time effects
fit_5a <- nnet::multinom(
  status ~ Gender + Age + Education_tri + Depression + Total_p + (Total_p * Age) +  status_prev + wave, family = multinomial, 
  data = data_stack, trace = FALSE)

fit_5b <- nnet::multinom(
  status ~ Gender + Age + Education_tri + Depression + Total_p + (Total_p * Age) + status_prev + wave, family = multinomial, 
  data = data_stack |> mutate(status = relevel(status, ref = 2)), 
  trace = FALSE)

# Model 6: State-specific time effects
fit_6a <- nnet::multinom(
  status ~ Gender + Age + Education_tri + Depression + Total_p + (Total_p * Age) + (status_prev * wave), family = multinomial, 
  data = data_stack, trace = FALSE)

fit_6b <- nnet::multinom(
  status ~ Gender + Age + Education_tri + Depression + Total_p + (Total_p * Age) + (status_prev * wave), family = multinomial, 
  data = data_stack |> mutate(status = relevel(status, ref = 2)), 
  trace = FALSE)

# Model 7: Full time interactions
fit_7a <- nnet::multinom(
  status ~ (Gender + Age + Education_tri + Depression + Total_p + (Total_p * Age) + status_prev) * wave, family = multinomial, 
  data = data_stack, trace = FALSE)

fit_7b <- nnet::multinom(
  status ~ (Gender + Age + Education_tri + Depression + Total_p + (Total_p * Age) + status_prev) * wave, family = multinomial, 
  data = data_stack |> mutate(status = relevel(status, ref = 2)), 
  trace = FALSE)
```

### Model comparison

Again, we will evaluate model improvement using likelihood ratio tests (@eq-liklihood; @tbl-non-stationary-comparison). However, this time, we will also compare the best fitted non-stationary models `fit_3`.

```{r}
#| code-fold: true
#| code-summary: "Check out my code"
#| label: tbl-non-stationary-comparison
#| tbl-cap: "Likelihood ratio test for non-stationary models"
#| tbl-cap-location: bottom

# Fit statistics
models <- list(
  "Interaction" = fit_4a,
  "Additive Time" = fit_5a,
  "State-Time Interaction" = fit_6a,
  "Full Interactions" = fit_7a
)

# Create comparison table
comparison_table <- purrr::map_dfr(models, broom::glance, .id = "Model") |>
  select(edf:AIC) |>
  mutate(across(where(is.numeric), \(x) round(x, digits = 1)))

# Outputting as table
anova(fit_4a, fit_5a, fit_6a, fit_7a) |>
  mutate(
    Model = c("Interaction", "Additive", "State-Time Interactions", "Full Interactions"),
    `Resid. Dev` = round(`Resid. Dev`, digits = 3),
    `LR stat.` = round(`LR stat.`, digits = 3),
    `Pr(Chi)` = round(`Pr(Chi)`, digits = 3)
    ) |>
  cbind(comparison_table) |>
  DT::datatable(
    options = list(
      pagelength = 4,
      dom = "t",
      ordering = FALSE,
      columnDefs = list(list(className = 'dt-center', targets = "_all"))
    ),
    rownames = FALSE
  )
```

**This is interesting!! Looks like adding in time as an additive term improved things. However, having full time interactions is better than having state-time interactions**

::: {.panel-tabset .nav-pills}
## Additive Model

### Estimated parameters

Key coefficients from the selected non-stationary model `fit_4`

```{r non-stationary-output}
additive_results <- rbind(tidy_output(fit_5a), tidy_output(fit_5b))

additive_results |>
  mutate(across(c(estimate, conf.low, conf.high), ~ round(x = ., digits = 3))) |>
  DT::datatable(
    options = list(
      pageLength = 6,
      ordering = FALSE,
      columnDefs = list(list(className = "dt-center", targets = "_all"))
    ),
    rownames = FALSE,
    colnames = c("Transition", "Predictor", "Estimate", "Standard Error",
                 "Statistic", "p-val", "Lower CI", "Upper CI")
  )
```

### Model visualisation

Let's visualize the predicted transition probabilities from the non-stationary model.

- @fig-non-stationary-age shows the predicted transition probabilities for age
- @fig-non-stationary-depression shows the predicted transition probabilities for depression
- @fig-non-stationary-education shows the predicted transition probabilities for per education level
- @fig-non-stationary-procrastination shows the predicted transition probabilities for procrastination

::: {.panel-tabset .nav-pills}

## Age

```{r}
#| code-fold: true
#| code-summary: "Check our my code"
#| fig-width: 10
#| fig-height: 8
#| label: fig-non-stationary-age
#| fig-cap: "Predicted transition probabilities from the non-stationary model (for age)."
#|

create_additive_predictions(data = data_stack, model = fit_5a, var = "Age", var_seq = seq(50, 97, by = 5)) |>
  plot_additive_predictions(var = "Age", x_label = "Age", subtitle = "By Age")

```

## Depression

```{r}
#| code-fold: true
#| code-summary: "Check our my code"
#| fig-width: 10
#| fig-height: 8
#| label: fig-non-stationary-depression
#| fig-cap: "Predicted transition probabilities from the non-stationary model (for depression)."
#|

create_additive_predictions(data = data_stack, model = fit_5a, var = "Depression", var_seq = seq(0, 8, by = 1)) |>
  plot_additive_predictions(var = "Depression", x_label = "Depression", subtitle = "By depression")

```

## Education Level

```{r}
#| code-fold: true
#| code-summary: "Check our my code"
#| fig-width: 10
#| fig-height: 8
#| label: fig-non-stationary-education
#| fig-cap: "Predicted transition probabilities from the non-stationary model (for having a high school degree)."

create_additive_predictions(data = data_stack, model = fit_5a, var = "Education_tri", var_seq = factor(c(0:2)), hold_constant = list(Gender = factor(0))) |>
  plot_additive_predictions(var = "Education_tri", x_label = "Education", subtitle = "By education level") +
  geom_point(size = 2) +
  scale_x_continuous(breaks = c(0, 1, 2), labels = c("No Education", "High School", "Further Education")) +
  theme(axis.text.x = element_text(size = 8, angle = 60, vjust = 0.90, hjust = 1))
```

## Procrastination

```{r}
#| code-fold: true
#| code-summary: "Check our my code"
#| fig-width: 10
#| fig-height: 8
#| label: fig-non-stationary-procrastination
#| fig-cap: "Predicted transition probabilities from the non-stationary model (for procrastination)."
#|

create_additive_predictions(data = data_stack, model = fit_5a, var = "Total_p", var_seq = seq(0, 60, by = 1)) |>
  plot_additive_predictions(var = "Total_p", x_label = "Procrastination", subtitle = "By procrastination")

```

## Interaction

```{r}
#| code-fold: true
#| code-summary: "Check our my code"
#| fig-width: 12
#| fig-height: 10
#| label: fig-non-stationary-interaction
#| fig-cap: "Predicted transition probabilities from the non-stationary model (interaction effect)."

# Making prediction dataset ----------------------------------------------------
pred_data <- expand.grid(
  Gender = factor(0),
  Age = c(62:97),
  Education_tri = factor(0),
  Depression = mean(data_stack$Depression),
  status_prev = levels(data_stack$status_prev),
  Total_p = seq(0, 60, length = 200),
  wave = c(2:4)
) |>
  modelr::add_predictions(model = fit_5a, var = "pred", type = "prob") |>
  tidy_predictions() |>
  filter(status_prev != "Dementia", status != "Dementia") |>
  mutate(label = case_when(
    Age %in% c(70, 80, 90) ~ as.character(Age), 
    TRUE ~ "Other"
  ))

# For adding age labels --------------------------------------------------------
label_data <- pred_data |>
  filter(Age %in% c(70, 80, 90)) |>
  group_by(status_prev, Age) |>
  filter(Total_p == max(Total_p)) |>
  ungroup()

pred_data |>
  ggplot(aes(x = Total_p, y = prob, 
             group = interaction(Age, status_prev, status),
             colour = label)) +
  geom_line(aes(linewidth = label)) +
  ggrepel::geom_text_repel(
    data = label_data,
    aes(label = paste0("Age ", Age), colour = as.character(Age)),
    direction = "y", nudge_x = 4, hjust = 0,
    segment.size = 1, segment.colour = "grey50",
    size = 2.5, show.legend = FALSE) +
  scale_x_continuous(
    breaks = seq(0, 60, by = 10), expand = expansion(mult = c(0, 0.2)),
    sec.axis = sec_axis( ~., name = "Previous State", breaks = NULL, labels = NULL)) +
  scale_y_continuous(sec.axis = sec_axis( ~ ., name = "Current State", breaks = NULL, labels = NULL)) +
  scale_colour_manual(values = c(
    "70"    = "#E69F00", 
    "80"    = "#56B4E9", 
    "90"    = "#009E73",
    "Other" = "grey80")) +
  scale_linewidth_manual(values = c(
    "70"    = 1.5, 
    "80"    = 1.5, 
    "90"    = 1.5,
    "Other" = 0.5)) +
  labs(
    title = "Predicted transition probabilities for different age cohorts",
    subtitle = "Across time",
    x = "Procrastination", y = "Probability", colour = NULL) +
  facet_grid(status ~ status_prev + wave, labeller = labeller(
    wave = c("2" = "2016 - 2018", "3" = "2018 - 2020", "4" = "2020 - 2022"),
  )) +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    strip.text = element_text(size = 9, face = "bold"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.title = element_text(size = 10, face = "bold"),
    legend.position = "none")

```

:::

### Estimated transition matrices

@fig-non-stationary-matrix-plot shows the estimated time-varying transition matrices

```{r}
#| code-fold: true
#| code-summary: "Check out my code"
#| fig-width: 12
#| label: fig-non-stationary-matrix-plot
#| fig-cap: "Estimated time-varying transition matrices."

# Create prediction grid for each time point
time_points <- unique(data_stack$wave)
state_names <- c("Normal Cognition", "MCI", "Dementia")

# Get matrices for all time points
time_varying_matrices <- purrr::map(
  setNames(time_points, paste(time_points, "Years")),
  ~ get_time_varying_matrix(fit_4a, .x)
)

# Converting into a tidy data frame
non_stationary_matrices <- purrr::imap_dfr(
  time_varying_matrices,
  ~ as.data.frame(.x) |> 
    tibble::rownames_to_column("From") |> 
    pivot_longer(-From, names_to = "To", values_to = "probability") |> 
    mutate(Time = .y),
  .id = "Time_point"
) |> 
  mutate(
    From = factor(From, levels = state_names),
    To = factor(To, levels = rev(state_names)),
    Time_point = forcats::fct_inorder(Time_point)
  )

# Plotting 
non_stationary_matrices |>
ggplot(aes(x = From, y = To, fill = probability)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(
    aes(label = format(round(probability, 3), nsmall = 3)),
    size = 4.5, 
    color = "#212427",
    fontface = "bold") +
    colorspace::scale_fill_continuous_diverging(
    palette = "Blue-Red 3", mid = 0.50, alpha = 0.5, 
    limits = c(0, 1), name = "Transition \nProbability") +
  labs(
    title = "Estimated Time-Varying Transition Matrices",
    subtitle = "Showing changes in transition probabilities over time",
    x = "Previous State (t-1)",
    y = "Next State (t)",
    fill = "Probability"
  ) +
  facet_wrap(~ Time_point, ncol = 3, labeller = labeller(
    Time_point = function(x){
      case_when(x == "2 Years" ~ "2018",
                x == "3 Years" ~ "2020",
                x == "4 Years" ~ "2022")
    })) +
  theme(
    axis.text = element_text(size = 10),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
    legend.position = "right",
    legend.text = element_text(size = 9),
    panel.grid = element_blank()
  )
```

## Multiplicative Model

### Estimated parameters

Key coefficients from the selected non-stationary model `fit_6`

```{r multiplicative-output}
multiplicative_output <- rbind(
  tidy_output(fit_7a, multiplicative = TRUE),
  tidy_output(fit_7b, multiplicative = TRUE))

multiplicative_output |>
  mutate(across(c(estimate, conf.low, conf.high), ~ round(x = ., digits = 3))) |>
  DT::datatable(
    options = list(
      pageLength = 6,
      ordering = FALSE,
      columnDefs = list(list(className = "dt-center", targets = "_all"))
    ),
    rownames = FALSE,
    colnames = c("Transition", "Predictor", "Estimate", "Standard Error",
                 "Statistic", "p-val", "Lower CI", "Upper CI", "colour")
  )
```

### Model visualisation

```{r}
multiplicative_output |>
  filter(!term %in% c("Previous state: MCI", "Previous state: Dementia", "Previous state: MCI & Time", "Previous state: Dementia & Time")) |>
  ggplot(aes(x = estimate, y = transition, colour = colour)) +
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey50") +
    ggstance::geom_pointrangeh(
    aes(xmin = conf.low, xmax = conf.high),
    position = ggstance::position_dodgev(height = 0.5),
    size = 1.25,
    fatten = 3) +
  scale_colour_manual(values = c(
    "Positive" = "#0072B2", 
    "Negative" = "#E69F00", 
    "NS"       = "#B2BEB5")) +
  facet_wrap(~term, scales = "free_x")
```

:::

# Matrix distance metrics

Given two matrices $P = (p_{ij})$ and $\hat{P} = (\hat{p}_{ij})$ of size $m \times n$, we define the following distance measures

**Frobenius Norm (Matrix Euclidean Distance)**

$$D_{\text{Frobenius}}(P, \hat{P}) = \|P - \hat{P}\|_F = \sqrt{\sum_{i=1}^{m}\sum_{j=1}^{n} |p_{ij} - \hat{p}_{ij}|^2}$$

- Measures the Euclidean distance in matrix space
- Sensitive to large differences due to squaring
- Equivalent to the vector 2-norm of the flattened matrix

**Manhattan Distance (L1 Norm)**

$$D_{\text{Manhattan}}(P, \hat{P}) = \sum_{i=1}^{m}\sum_{j=1}^{n} |p_{ij} - \hat{p}_{ij}|$$

- Sum of absolute differences between corresponding elements
- Less sensitive to outliers than Frobenius norm
- Useful when sparse differences are important

**Maximum Difference (Infinity Norm)**

$$D_{\text{Max}}(A, B) = \max_{\substack{1 \leq i \leq m \\ 1 \leq j \leq n}} |a_{ij} - b_{ij}|$$

- Captures the single largest difference between elements
- Useful for worst-case analysis
- Ignores the distribution of other differences

**Mean Absolute Difference**

$$D_{\text{MeanAbs}}(P, \hat{P}) = \frac{1}{mn}\sum_{i=1}^{m}\sum_{j=1}^{n} |p_{ij} - \hat{p}_{ij}|$$

- Average absolute difference across all elements
- Scales with matrix size (unlike Manhattan distance)
- Easy to interpret as average error per element

**Root Mean Square Error (RMSE)**

$$D_{\text{RMSE}}(P, \hat{P}) = \sqrt{\frac{1}{mn}\sum_{i=1}^{m}\sum_{j=1}^{n} (p_{ij} - \hat{p}_{ij})^2}$$

- Similar to Frobenius but normalized by matrix size
-  Sensitive to large errors due to squaring
- Common in statistical and machine learning applications

**Correlation-Based Distance**

$$D_{\text{Corr}}(P, \hat{P}) = 1 - \frac{\sum_{i=1}^{m}\sum_{j=1}^{n} (p_{ij} - \bar{P})(\hat{p}_{ij} - \bar{\hat{P}})}{\sqrt{\sum_{i=1}^{m}\sum_{j=1}^{n} (p_{ij} - \bar{P})^2 \sum_{i=1}^{m}\sum_{j=1}^{n} (\hat{p}_{ij} - \bar{\hat{p}})^2}}$$

- Measures linear relationship between matrix elements
- $\bar{P} = \frac{1}{mn}\sum_{i,j}p_{ij}$ (mean of all elements in $P$)
- $\bar{\hat{P}} = \frac{1}{mn}\sum_{i,j}\hat{p}_{ij}$ (mean of all elements in $\hat{P}$)
- Range: [0,2] where 0 = perfect positive correlation

**Kullback-Leibler Divergence**

For matrices where each row sums to 1 ($\sum_j p_{ij} = \sum_j \hat{P}_{ij} = 1$) and $p_{ij}, \hat{p}_{ij} > 0$:

$$
D_{\text{KL}}(P \parallel \hat{P}) = \sum_{i=1}^{m}\sum_{j=1}^{n} p_{ij} \log\left(\frac{p_{ij}}{\hat{p}_{ij}}\right)
$$

- Measures information loss when $\hat{P}$ approximates $P$
- Asymmetric: $D_{\text{KL}}(P \parallel \hat{P}) \neq D_{\text{KL}}(\hat{P} \parallel P)$
- $0\log0 = 0$ by convention
- In practice, add $\epsilon > 0$ (e.g., $10^{-10}$) to avoid zeros

## Implementation

Let's calculate the distance metrics between the observed $P$ and estimated transition matrices $\hat{P}$.

```{r matrix-distance}
p       <- transition_matrix_observed
p_hat   <- transition_matrix_estimated
epsilon <- 1e-10

# Creating tibble of distance metrics
distances <- tibble(
  Metric = c("Frobenius", "Manhattan", "Max", "MeanAbs", "RMSE", "Correlation", "KL"),
  Value = c(
    norm(p - p_hat, type = "F"),
    sum(abs(p - p_hat)),
    max(abs(p - p_hat)),
    mean(abs(p - p_hat)),
    sqrt(mean((p - p_hat)^2)),
    1 - cor(c(p), c(p_hat)),
    sum((p + epsilon) * log((p + epsilon) / (p_hat + epsilon)))
  )) |>
  mutate(Value = round(Value, 4))

distances |>
  mutate(Metric = case_when(
    Metric == "Frobenius" ~ "Frobenius Distance",
    Metric == "Manhattan" ~ "Manhattan Distance",
    Metric == "Max" ~ "Max Difference",
    Metric == "MeanAbs" ~ "Mean Absolute Difference",
    Metric == "RMSE" ~ "Root Mean Square Error",
    Metric == "Correlation" ~ "Correlation Distance",
    Metric == "KL" ~ "Kullback-Leibler Divergence"
  )) |>
  knitr::kable(caption = "Distance based metrics", align = "c") |>
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = FALSE)

```

## Visualisation

Let's visualize the distance metrics between the observed and estimated transition matrices (@fig-metrics).

```{r}
#| code-fold: true
#| code-summary: "Check out my code"
#| fig-width: 12
#| label: fig-metrics
#| fig-cap: "Distance metrics between observed and estimated transition matrices."

caption = stringr::str_glue(
  "**RMSE:** Root Mean Squared Error\n
   **KL:** Kullback-Leibler Divergence\n
   **Correlation:** 1 - Pearson Correlation Coefficient")

distances |>
  ggplot(aes(x = reorder(Metric, -Value), y = Value, fill = Metric)) +
  geom_col(colour = "black") +
  geom_text(aes(label = Value), vjust = -0.5) +
  ggokabeito::scale_fill_okabe_ito() +
  scale_y_continuous(expand = expansion(mult = c(0.075, 0.075))) +
  labs(
    title = "Distance Between Transition Matrices",
    x = "Distance Metric",
    y = "Value",
    caption = caption) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(face = "bold", size = 12),
    plot.caption = ggtext::element_markdown(size = 10),
    legend.position = "none")
```

```{r test-code}
#| include: false
#| eval: false
preds <- create_additive_predictions(data = data_stack, model = fit_4a, var = "Age", var_seq = seq(50, 97, by = 5)) |>
  # Add color column based on transition significance
  mutate(
    transition_type = paste(status_prev, status, sep = " → "),
    colour = case_when(
      transition_type == "Normal Cognition → Normal Cognition" ~ "#E69F00",
      transition_type == "Normal Cognition → MCI" ~ "#E69F00",
      transition_type == "Normal Cognition → Dementia" ~ "#B2BEB5",
      transition_type == "MCI → Normal Cognition" ~ "#56B4E9",
      transition_type == "MCI → MCI" ~ "#56B4E9",
      transition_type == "MCI → Dementia" ~ "darkgrey",
      transition_type == "Dementia → Normal Cognition" ~ "#009E73",
      transition_type == "Dementia → MCI" ~ "#009E73",
      transition_type == "Dementia → Dementia" ~ "#009E73"
    ))

ggplot(preds, aes(Age, prob, color = colour, group = interaction(wave, transition_type))) +
  geom_line(linewidth = 1) +
  scale_color_identity("Previous State", 
                      breaks = c("#E69F00", "#56B4E9", "#009E73"),
                      labels = c("Normal Cognition", "MCI", "Dementia"),
                      guide = "legend") +
  labs(
    title = "Transition Probabilities (Additive Model)",
    subtitle = "By Age",
    x = "Age", 
    y = "Predicted Probability",
    caption = "*Note:* Grey lines indicate non-significant transitions",
    ) +
    facet_grid(wave ~ status, labeller = labeller(
      wave = function(x) case_when(
        x == 2 ~ "Year = 2018",
        x == 3 ~ "Year = 2020",
        x == 4 ~ "Year = 2022",
      ),
      status = function(x) paste("Transition to:", x)
    )) +
    theme(
      plot.subtitle = element_text(hjust = 0.5, size = 12, face = "bold"),
      plot.caption = ggtext::element_markdown(),
      panel.spacing = unit(1, "lines"))

```