### 2025-09-30 README.qmd
Adding scripts, files, and manuscript :fire:

---
title: "Procrastination as a Marker of Cognitive Decline: Evidence from Longitudinal Transitions in the Older Adult Population"
format: gfm
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r setup, include=FALSE}
library(emoji)

yaml_metadata <- rmarkdown::yaml_front_matter(here::here("manuscript", "manuscript.qmd"))

title <- yaml_metadata$title
abstract <- yaml_metadata$abstract

emoji_out <- emoji_glue(":potato::spiral_shell:")
```

::: {.content-visible when-format="markdown"}

---

- [Cormac Monaghan](https://c-monaghan.github.io/) • Maynooth University
- [Rafael de Andrade Moral](https://rafamoral.github.io/) • Maynooth University
- [Michelle Kelly](https://www.ncirl.ie/About/A-Z-Staff-Directory/Staff/346) • National College of Ireland
- [Joanna McHugh Power](https://www.maynoothuniversity.ie/people/joanna-mchugh-power) • Maynooth University

---

:::

---

> Monaghan, C., de Andrade Moral, R, Kelly, M, & McHugh Power, J. "`r title`".

---

## Abstract

`r abstract`

## `r emoji_out`: Note on "crisp clam" project name

Because project titles change all the time we used [{codename}](http://svmiller.com/codename/) to generate an [Ubuntu-style](https://wiki.ubuntu.com/DevelopmentCodeNames)
internal-to-us project name that won’t change.

```{r code-name, eval=FALSE}
library(codename)

codename_message()
#> code name generated by {codename} v.0.5.0. R version 4.4.2 (2024-10-31 ucrt).

codename(seed = "dementia", type = "ubuntu")
#> [1] "crisp clam"
```