### 2025-09-30 analysis/discrete-markov.qmd

---
title: "Discrete Time Markov Model"
---

# Setting up

```{r}
#| label: setup
#| warning: false
#| message: false
#| output: false
#| code-fold: true
#| code-summary: "Check out my code"

rm(list = ls())

# Packages ---------------------------------------------------------------------
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)

# Functions --------------------------------------------------------------------
files <- list.files(here::here("R/"), full.names = TRUE)

purrr::walk(files, source)

# Theme ------------------------------------------------------------------------
colour <- "#212427"

theme_set(
  theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, size = 14, colour = colour, face = "bold"),
      plot.subtitle = element_text(hjust = 0.5, size = 12, colour = colour),
      axis.title = element_text(size = 10, colour = colour, face = "bold"),
      strip.text = element_text(size = 10, colour = colour, face = "bold"),
      legend.title = element_text(hjust = 0.5, colour = colour, face = "bold"),
      ))

# Data -------------------------------------------------------------------------
data <- read.csv(here::here("analysis/data/data.csv"))
```

# Data Preparation

Before preparing the data, we want to create a vector of covariates that we are interested in using in our analysis

```{r}
#| label: covariates
#| 
cols <- c(
  "ID", "Gender", "Age", "Education_tri",
  paste0("apathy_", seq(2016, 2022, by = 2)),
  "Total_p")
```

Following this, we can now use the `extract_years()` function to extract the cognitive function data for the years 2016, 2018, and 2020.

- We will also impute some missing values `impute = TRUE` using logical reasoning (if a respondent has an NA value in 2018, but has a classification of "normal cognition" in 2020, then the missing 2018 value becomes "normal cognition").
- We will also treat dementia as an absorbing state `absorbing = TRUE`.

```{r}
#| label: extract-years
#| collapse: true
data_stack <- data |> 
  extract_years(years = seq(2016, 2022, by = 2), impute = TRUE, absorbing = TRUE) |>
  na.omit()

head(data_stack)
```

## Creating a stacked dataset

For each respondent we will now add in their relevant covariate data. Following this, we transform the data to long format and convert categorical variables to factors using the `pivot_and_factorise()` function. Additionally, we will fix the `Age` column to properly represent the age of the respondent at each time point

```{r}
#| label: long_data
#| collapse: true
data_stack <- data_stack |>
  # Join covariate data
  inner_join(data[, cols], by = "ID") |>
  # Removing baseline age below 60
  filter(Age - (2022 - 2016) >= 60) |>
  pivot_and_factorise(time_vary = TRUE) |>
  group_by(ID) |>
  # Make age time varying
  mutate(Age = Age - (2022 - as.numeric(as.character(wave))))

head(data_stack)
```

Finally, we will create a new variable `status_prev` that notes the cognitive status of the respondent in the previous wave $(t - 1)$. This will be done by using the `lag` function from the `dplyr` package.

```{r}
#| label: stack-data
#| collapse: true
#| 
data_stack <- data_stack |>
  mutate(status_prev = lag(status), .after = status) |>
  filter(!is.na(Total_p)) |>
  ungroup() |> 
  filter(wave != 2016)

head(data_stack)
```

# Modelling
## Markov Process Fundamentals

Discrete-time Markov models belong to a class of stochastic processes that satisfy the **Markov property**, which can be formally expressed as:

$$
P(X_{t+1} = j \vert X_t = i, X_{t-1} = i_{t-1}, \dots X_0 = i_0) = P(X_{t+1} = j \vert X_t = i)
$$ {#eq-markov}

This property establishes that the future state $X_{t+1}$ depends only on the current state $X_t$, not on the entire history of states.

### Multinomial Logistic Regression Formulation

We model the transition probabilities using **multinomial logistic regression**, where the log-odds of each transition relative to a reference state are linear functions of covariates.

For a system with $K$ states (using state $K$ as reference), we have:

$$
log \left( \frac{P(Y = j \vert x)}{P(Y = k \vert x)} \right) = \beta_0 + \beta_j^Tx \qquad \text{for } j = 1, \dots K-1
$$ {#eq-multinomial}

For non-reference states $j = 1, \dots, K - 1$

$$
P(Y = j \vert x) = \frac{e^{\beta_{0j} + \beta_j^Tx}}{1 + \sum^{k - 1}_{k = 1} e^{\beta_{0k} + \beta_k^Tx}}
$$ {#eq-non-reference}

For the reference state $K$:

$$
P(Y = k \vert x) = \frac{1}{1 + \sum^{k - 1}_{k = 1} e^{\beta_{0k} + \beta_j^Tx }}
$$ {#eq-reference}

### Absorbing State Specification

Our primary analysis assumes the transition matrix remains fixed:

$$
P^{(t)} \equiv P \qquad \forall \; t
$$ {#eq-assumption}

However, we model Dementia as an absorbing state:

$$
p_{3j} = \begin{cases} 
1 \qquad \text{if } j = 3 \\
0 \qquad \text{otherwise} \end{cases}
$$ {#eq-dementia}

yielding the constrained transition matrix:

$$
P = \begin{bmatrix} 
p_{11} & p_{12} & p_{13} \\
p_{21} & p_{22} & p_{23} \\
0      & 0      & 1      \\
\end{bmatrix}
$$ {#eq-t-matrix-constrained}

## Our model

As mentioned, we estimate our Markov model using `nnet::multinom()`

```{r}
#| label: model-fit

fit_a <- nnet::multinom(
  status ~ Gender + Education_tri + Depression + (Total_p * Age) + status_prev, 
  family = multinomial, 
  data = data_stack, trace = FALSE)

fit_b <- nnet::multinom(
  status ~ Gender + Education_tri + Depression + (Total_p * Age) + status_prev, 
  family = multinomial, 
  data = data_stack |> mutate(status = relevel(status, ref = 2)), 
  trace = FALSE)
```

### Estimated parameters

Below we present the estimated parameters for the interaction model (@tbl-interaction-output). We will transform the estimates to odds ratios and colour them based on significance.

```{r}
#| label: tbl-interaction-output
#| tbl-cap: "Estimated parameters for the interaction model"
#| tbl-cap-location: bottom
#| code-fold: true
#| code-summary: "Check out my code"

interaction_results <- rbind(tidy_output(fit_a), tidy_output(fit_b)) |>
  mutate(
    # Mutating to factors
    transition = factor(
      transition, 
      levels = c("NC - MCI", "MCI - NC", 
                 "NC - Dementia", "MCI - Dementia")),
    term = factor(
      term,
      levels = c("Being female", "Age", 
                 "High school degree vs. No education",
                 "Further education vs. No education", "Depression",
                 "Procrastination (2020)",  
                 "Age x Procrastination", "Previous state: MCI", 
                 "Previous state: Dementia")),
    )

interaction_results |>
  mutate(across(c(estimate, conf.low, conf.high), ~ round(x = ., digits = 4))) |>
  DT::datatable(
    options = list(
      pageLength = 10,
      dom = "tip",
      ordering = FALSE,
      columnDefs = list(list(className = "dt-center", targets = "_all"))
    ),
    rownames = FALSE
  )
```

### Model visualisation

Let’s visualize the estimated odds ratios (@fig-odds-ratio-interaction) and predicted transition probabilities (@fig-interaction-predictions) for our interaction model.

```{r}
#| label: fig-odds-ratio-interaction
#| fig-width: 12
#| fig-height: 10
#| code-fold: true
#| code-summary: "Check out my code"
#| fig-cap: "Estimated odds ratios for the interaction model."
#| message: false
#| warning: false

# Design matrix for faceting
design <- "
  AABBCC 
  DDEEFF
  ##GG##
"

interaction_results |>
  filter(!term %in% c("Previous state: MCI", "Previous state: Dementia")) |>
  # Adding text labels
  mutate(
    label_pos = case_when(
      stringr::str_detect(term, "Further") & 
        stringr::str_detect(transition, "MCI - NC") ~ estimate - 0.5,
      stringr::str_detect(term, "Further") ~ estimate - 0.5,
      TRUE ~ estimate
    ),
    label_text = sprintf(
    "<span style='line-height:3.5'>OR = %.3f<br><br>[%.3f, %.3f]</span>",
    estimate, conf.low, conf.high)) |>
  # Plotting
  ggplot(aes(x = estimate, y = transition, colour = colour)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray50") +
  # Odds ratios
  ggstance::geom_pointrangeh(
    aes(xmin = conf.low, xmax = conf.high),
    position = ggstance::position_dodgev(height = 0.6),
    size = 1,
    fatten = 2) +
  # Estimates and CIS
  ggtext::geom_richtext(
    aes(x = label_pos, label = label_text),
    position = ggstance::position_dodge2v(height = 0.5),
    hjust = 0,
    size = 3,
    fill = NA, label.colour = NA,
    show.legend = FALSE) +
  # Customization
  scale_colour_manual(values = c(
    "Positive" = "#0072B2", 
    "Negative" = "#E69F00", 
    "NS"       = "#B2BEB5")) +
  # scale_x_continuous(expand = expansion(mult = 0.025, add = 0.025)) +
  labs(# title = "Odds ratios (stationary model)",
       x = "Odds Ratio", y = "Predictor") +
  guides(colour = "none") +
  ggh4x::facet_manual(
    ~ term, scales = "free_x", 
    design = design, labeller = labeller(
     term = c("Depression" = "Apathy")
    )) +
  coord_cartesian(clip = "off") +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    strip.text = element_text(size = 10, face = "bold"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank())

```

```{r}
#| label: fig-interaction-predictions
#| fig-width: 12
#| fig-height: 10
#| code-fold: true
#| code-summary: "Check out my code"
#| fig-cap: "Predicted transition probabilities from the interaction model."
#| warning: false

# Making a new predicted dataset
pred_data <- expand.grid(
  Gender = factor(0),
  Age = seq(60, 97, length = 200),
  Education_tri = factor(0),
  Depression = mean(data_stack$Depression, na.rm = TRUE),
  status_prev = levels(data_stack$status_prev),
  Total_p = seq(0, 60, length = 200))

# Reveling original data
data_plot <- data_stack |>
  mutate(
    # Relevel
    status = case_when(
      status == 1 ~ "Normal Cognition",
      status == 2 ~ "MCI",
      status == 3 ~ "Dementia"),
    status_prev = case_when(
      status_prev == 1 ~ "Normal Cognition",
      status_prev == 2 ~ "MCI",
      status_prev == 3 ~ "Dementia")
    ) |>
  mutate(
    status = factor(
      status, levels = c("Normal Cognition", "MCI", "Dementia")),
    status_prev = factor(
      status_prev, levels = c("Normal Cognition", "MCI", "Dementia"))) |>
  filter(status_prev %in% c("Normal Cognition", "MCI") & 
         status %in%  c("Normal Cognition", "MCI"))

# Plotting interaction effect ----------------------------------------------
pred_data |>
  # Adding predictions
  modelr::add_predictions(model = fit_a, var = "pred", type = "probs") |>
  tidy_predictions() |>
  filter(status_prev %in% c("Normal Cognition", "MCI") & 
         status %in%  c("Normal Cognition", "MCI")) |>
  ggplot(aes(x = Total_p, y = as.numeric(Age))) +
  # Adding heat mat with contour lines
  geom_raster(aes(fill = prob), alpha = 0.7) +
  geom_contour(aes(z = prob), colour = "black", size = 0.3) +
  # Adding orginal data points
  geom_point(
    data = data_plot, aes(x = Total_p, y = Age), size = 1.75, alpha = 0.3) +
  # Customising plot
  scale_x_continuous(
    breaks = seq(0, 60, by = 10),
    sec.axis = sec_axis(~ ., name = "Previous State",
                        breaks = NULL, labels = NULL)) +
  scale_y_continuous(
    breaks = seq(50, 100, by = 5),
    sec.axis = sec_axis(~ ., name = "Current State",
                        breaks = NULL, labels = NULL)) +
  scale_fill_viridis_c(option = "plasma") +
  labs(
    title = "Predicted transition probabilities from the interaction model",
    x = "Total Procrastination", y = "Age", 
    fill = expression(hat(p))) +
  facet_grid(status ~ status_prev) +
  theme_classic() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    strip.text = element_text(size = 10, face = "bold"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.title = element_text(size = 10, face = "bold"))

```

Finally, (@fig-interaction-age-predictions) shows the the predicted transition probabilities for different age cohorts.

```{r}
#| label: fig-interaction-age-predictions
#| fig-width: 12
#| fig-height: 8
#| code-fold: true
#| code-summary: "Check out my code"
#| fig-cap: "Predicted transition probabilities for different age cohorts."
#| warning: false

# Create predictions with age info
age_cohots <- expand.grid(
  Gender = factor(0),
  Age = c(62:97),
  Education_tri = factor(0),
  Depression = mean(data_stack$Depression, na.rm = TRUE),
  status_prev = levels(data_stack$status_prev),
  Total_p = seq(0, 60, length = 200)) |>
  modelr::add_predictions(model = fit_a, var = "pred", type = "probs") |>
  tidy_predictions() |>
  filter(status_prev %in% c("Normal Cognition", "MCI") &
         status %in%  c("Normal Cognition", "MCI")) |>
  mutate(label = case_when(
    Age %in% c(70, 80, 90) ~ as.character(Age), 
    TRUE ~ "Other"
  ))

# Identifying right most point for each label point
label_data <- age_cohots |>
  filter(Age %in% c(70, 80, 90)) |>
  group_by(status_prev, Age) |>
  filter(Total_p == max(Total_p)) |>
  ungroup()

# Plotting ---------------------------------------------------------------------
ggplot(data = age_cohots, aes(x = Total_p, y = prob, 
                              group = interaction(Age, status_prev, status),
                              colour = label)) +
  geom_line(aes(linewidth = label)) + 
  ggrepel::geom_text_repel(
    data = label_data,
    aes(label = paste0("Age ", Age), colour = as.character(Age)),
    direction = "y", nudge_x = 4, hjust = 0,
    segment.size = 0.5, segment.colour = "grey50",
    size = 3, show.legend = FALSE) +
  scale_x_continuous(
    breaks = seq(0, 60, by = 10),
    sec.axis = sec_axis( ~., name = "Previous State", breaks = NULL, labels = NULL)) +
  scale_y_continuous(sec.axis = sec_axis( ~ ., name = "Current State", breaks = NULL, labels = NULL)) +
  scale_colour_manual(values = c(
    "70"    = "#E69F00", 
    "80"    = "#56B4E9", 
    "90"    = "#009E73",
    "Other" = "grey80")) +
  scale_linewidth_manual(values = c(
    "70"    = 1.5, 
    "80"    = 1.5, 
    "90"    = 1.5,
    "Other" = 0.5)) +
  labs(
    # title = "Predicted transition probabilities for different age cohorts",
    x = "Procrastination", y = "Probability", colour = NULL) +
  facet_grid(status ~ status_prev, labeller = labeller(
    status = c("Normal Cognition" = "Normative Cognitive Function"),
    status_prev = c("Normal Cognition" = "Normative Cognitive Function")
  )) +
  theme_bw() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.title = element_text(size = 12, face = "bold"),
    strip.text = element_text(size = 10, face = "bold"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.title = element_text(size = 10, face = "bold"),
    legend.position = "none")
```