### 2025-11-25 analysis/hypothesis.qmd

---
title: "Hypotheses"
format:
  html:
    code-tools: false
    toc: false
---

This study was **not pre-registered** on the [Open Science Framework](https://osf.io/). However, we had two main hypotheses.

## Hypothesis 1

We hypothesize that:

- **H0:** There are no group differences in procrastination scores between cognitive groups (normative cognitive function; MCI; Dementia).
- **H1:** There are group differences in procrastination scores between cognitive groups (normative cognitive function; MCI; Dementia)

## Hypothesis 2

We hypothesize that:

- **H0:** There is no association between procrastination and cognitive decline transitions
- **H1:** Higher scores on a measure of procrastination are associated with:
  - Increased risk of transitioning from normative cognitive function to MCI / dementia $(\beta = x, p < .05)$
  - Increased risk of transitioning from MCI to Dementia $(\beta = y, p < .05)$

In order to test our hypotheses we will be using both a **non-parametric ANOVA** and a post-hoc test to compare procrastination scores between cognitive groups, and a **discrete time Markov Model** to assess the association between procrastination and cognitive decline transitions (controlling for known covariates).